### 2014-04-20 paper/paper.tex
Add paper stuff

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}

\title{
Probabilistic Programming
%\\\vspace{4 mm}\underline{\Large House Lannister}
}

\author{
	Carlos Diaz
	\and
	Dan Hassin
	\and
	Charles Lehner
	\and
	Dan Viterise
}

\date{April 21, 2014}

\begin{document}

\maketitle

%\begin{abstract}
	%Your abstract goes here...
	%...
%\end{abstract}

\section{Introduction}

\subsection{Motivation}

% overview of markov chain
% application of markov chain to abstact syntrax tree

\section{Implementation}

\end{document}

### 2014-04-20 generate.js
Fix ExpressionStatement bug

@@ -68,8 +68,9 @@ function generateList(model, path)
     }
 
     while (node.type != END) {
+        var type = (node.type == "ExpressionStatement" ? node.expression.type : node.type);
         nodes.push(node);
-        path = path.concat(node.type, FOLLOW);
+        path = path.concat(type, FOLLOW);
         node = generateNode(model, path);
     }
     return nodes;
@@ -89,6 +90,8 @@ function generateNode(model, path, sel, depth)
     //cut off the path at given depth
     var maxPathLength = -(DEPTH)*2;
     path = path.slice(maxPathLength);
+
+    // console.log("generating with path "+path);
 
     var prevTransition = path[path.length-1];
     var prevNodeType = path[path.length-2];
@@ -203,6 +206,14 @@ function instantiateNode(type, model, path)
     {
         //turn raw into a value, needed by escodegen, with eval
         node["value"] = eval(node.raw);
+
+        var my_escape = function (string) {
+            if (typeof string != "string") return string;
+            return string.replace(/>/g,"&gt;").replace(/</g,"&lt;").replace(/&/g,"&amp;");
+        };
+
+        if (typeof node.value == "string")
+            node.value = my_escape(node.value);
     }
 
     return node;

### 2014-04-20 parse.js

@@ -76,6 +76,7 @@ function addCount(node, path)
 	var key = (node.type == END) ? type : type.toString().replace(/_/g, "__");
 	probs[key] = (Object.prototype.hasOwnProperty.call(probs, key) ? probs[key] : 0) + 1;
 	probs[TOTAL] += 1;
+	//console.log("addCount for ",key, "path = "+ path);
 	if (node.expr) {
 		// mark that the node is wrapped by an expression
 		var expr = probs['_expr'] || (probs['_expr'] = {});
@@ -88,6 +89,9 @@ function parseList(nodes, path)
 	for (var i = 0; i < nodes.length; i++) {
 		var node = nodes[i];
 		parseNode(node, path);
+		if (node.type == "ExpressionStatement") {
+			node = node.expression;
+		}
 		path = path.concat(node.type, FOLLOW);
 	}
 	parseNode(END_NODE, path);

### 2014-04-20 rand.js

@@ -20,7 +20,7 @@ fs.readFile("parse.js", 'utf8', function (err,data) {
         {
             //print out the model & syntax if an extra param is passed in
             console.log(JSON.stringify(model, null, 4));
-            console.log(JSON.stringify(syntax, null, 4));
+            //console.log(JSON.stringify(syntax, null, 4));
         }
         else
         {

### 2014-04-20 web/main.js

@@ -38,10 +38,15 @@ function getPath(syntax, node, depth, path, container, idx) {
 	if (!syntax || typeof syntax != "object") {
 		return;
 	}
+	if (syntax.type == "ExpressionStatement")
+		syntax = syntax.expression;
+
 	if (objectsEqual(syntax, node)) {
-		return {"path":path.concat(node.type, FOLLOW),
-			"container":container, "idx":idx};
+		if (path[path.length-1] == FOLLOW || path[path.length-1] == 'body')
+			path = path.concat(node.type, FOLLOW)
+		return {"path":path, "container":container, "idx":idx};
 	}
+
 	if (syntax.type) {
 		path = path.concat(syntax.type);
 		if (depth) {
@@ -52,6 +57,9 @@ function getPath(syntax, node, depth, path, container, idx) {
 		for (var i = 0; i < syntax.length; i++) {
 			var item = syntax[i];
 			if (!item) continue;
+			if (item.type == "ExpressionStatement")
+				item = item.expression;
+
 			var p = getPath(item, node, depth, path, syntax, i);
 			if (p) return p;
 			path.push(item.type, FOLLOW);
@@ -71,6 +79,9 @@ function setCurrentPath(path) {
 }
 
 window.setCurrentPathToNode = function (node) {
+
+	if (node.type == "ExpressionStatement")
+		node = node.expression;
 
 	console.log("looking for", node, " in ",tree);
 	currentPosition = getPath(tree, node);
@@ -101,15 +112,15 @@ function generateProgramSource() {
 	console.log("**** generating snippet from path ",currentPosition.path);
 
 	var syntax = generate.generateNode(model, currentPosition.path, true, 2);
-	if (syntax.type == END) {
+	if (!syntax || syntax.type == END) {
 		return;
 	}
 	var source;
-	try {
+	//try {
 		source = escodegen.generate(syntax, escodegenOptions);
-	} catch(e) {
-		console.log("generate", e);
-	}
+	// } catch(e) {
+	// 	console.log("generate", e);
+	// }
 	return source;
 }
 

### 2014-04-20 web/ui.js

@@ -67,7 +67,6 @@ function complete(option)
 	if (node.type != "Program")
 		window.setCurrentPathToNode(node);
 
-
 	var snippet = $(option).children();
 	snippet.insertAfter(cursor);
 	$("<br>").insertAfter(cursor);
@@ -93,9 +92,9 @@ function newOptions()
 		$('#options').append($('<li></li>').append(pre));
 	}
 
-	$('#options').append($('<li><pre contenteditable="true"></pre></li>'));
+	//$('#options').append($('<li><pre contenteditable="true"></pre></li>'));
 
-	numSelections = snips.length+1;
+	numSelections = snips.length;
 
 	currentSelection = 0;
     $($('#options').children()[currentSelection]).find('pre').css('background-color','#D7EBFC');

### 2014-04-20 web/ui.js
Ignore block statements

@@ -34,8 +34,8 @@ function checkKey(e)
 
 function isStatement(nodeType)
 {
-	return nodeType == "VariableDeclaration" ||
-		nodeType.match(/Statement$/);
+	if (nodeType == "BlockStatement") return false;
+	return nodeType.match(/(Statement|Declaration)$/);
 }
 
 var cursor;
@@ -84,7 +84,7 @@ function newOptions()
 {
 	$('#options').html("");
 
-	var snips = window.getSnippets(3);
+	var snips = window.getSnippets(6);
 	for (var i = 0; i < snips.length; i++)
 	{
 		code = snips[i];

### 2014-04-20 web/ui.js
Handle clicks on snippets

@@ -40,6 +40,12 @@ function isStatement(nodeType)
 
 var cursor;
 function addHandlers() {
+	$('#options').on('click', 'pre', function() {
+		var options = $('#options').children();
+		$(options[currentSelection]).find('pre').css('background-color','clear');
+		currentSelection = $(this.parentNode).index();
+		$(options[currentSelection]).find('pre').css('background-color','#D7EBFC');
+	});
 	$("#editor").on('click', '.statement', function(e) {
 		var el = $(this);
 		var node = el.data('node');

### 2014-04-20 paper/paper.tex
Update paper.tex
Began Introduction

@@
 \section{Introduction}
-
+For many people just starting out, programming can seem like a confusing and tedious activity. It can be unclear for 
+many new programmers where they should begin and what should come next while coding. Without any proper guidance, 
+a novice programmer is left on their own to scour thousands of pages online to discover what should come next in 
+their code. Then, once a person becomes more adept at programming, he or she may notice that they are typing out 
+the same or similar lines of code within each new program. This repetitiveness slows down the user and can make 
+programming a much more tiresome task than it needs to be. 
 \subsection{Motivation}

### 2014-04-20 web/main.js
Indentation

@@ -83,13 +83,13 @@ window.setCurrentPathToNode = function (node) {
 	if (node.type == "ExpressionStatement")
 		node = node.expression;
 
-	console.log("looking for", node, " in ",tree);
+	// console.log("looking for", node, " in ",tree);
 	currentPosition = getPath(tree, node);
 	if (!currentPosition) {
 		console.log("tree:", JSON.stringify(tree), 'Node:', JSON.stringify(node));
 	}
 
-	console.log("current position is ",currentPosition);
+	// console.log("current position is ",currentPosition);
 }
 
 window.insertSnippet = function (node) {

### 2014-04-20 web/ui.js

@@ -55,9 +55,19 @@ function addHandlers() {
 
 function complete(option)
 {
-	console.log("generating from position ",window.getCurrentPosition());
+	var selectedLine = cursor.parent().text();
+	selectedLine = selectedLine.replace(/^\{\n/, "");
+	var spaces = "";
+	var numspaces = selectedLine.match(/^ */)[0].length;
+	if (numspaces == 1) numspaces = 0;
+	for (var i = 0; i < numspaces; i++)
+		spaces += " ";
+	console.log("line = ",selectedLine,"  spaces = ",numspaces);
 
-	console.log("cursor");
+
+	// console.log("generating from position ",window.getCurrentPosition());
+
+	// console.log("cursor");
 
 	var node = $($(option).children()[0]).data('node');
 	console.log("snippet = ",$($(option).children()[0])," node = ",node);
@@ -66,9 +76,14 @@ function complete(option)
 
 	if (node.type != "Program")
 		window.setCurrentPathToNode(node);
-
+	
 	var snippet = $(option).children();
-	snippet.insertAfter(cursor);
+	snippet.each(function () {
+		if (numspaces > 0)
+			//$(this).prepend($("<span>"+spaces+"</span>"));
+			$(this).css('padding-left','16px').css('display','inline-block');
+		$(this).insertAfter(cursor);
+	});
 	$("<br>").insertAfter(cursor);
 	newOptions();
 

### 2014-04-20 paper/paper.tex
Update paper.tex
Finished Introduction (early draft)

@@
 the same or similar lines of code within each new program. This repetitiveness slows down the user and can make 
-programming a much more tiresome task than it needs to be. 
+programming a much more tiresome task than it needs to be. These observations steered our group towards developing
+a way for someone to code with guidance. The Probabilistic Programming method allows for a user to code in JavaScript
+using a web application. This web app provides a list of suggestions for the user's next line of code depending on where
+the user is in the program. Not only is this helpful for new programmers that need suggestions for what to write next 
+in their code, this system allows for expert programmers to write code with a great deal more ease and efficiency.
 \subsection{Motivation}

### 2014-04-20 paper/paper.tex
Update paper.tex
Added Markov Chain description under Motivation
Added References Section
Made small changes to Introduction

@@
 programming a much more tiresome task than it needs to be. These observations steered our group towards developing
-a way for someone to code with guidance. The Probabilistic Programming method allows for a user to code in JavaScript
+a way for someone to code with more instruction and productivity. The Probabilistic Programming method allows for a user to code in JavaScript
 using a web application. This web app provides a list of suggestions for the user's next line of code depending on where
-the user is in the program. Not only is this helpful for new programmers that need suggestions for what to write next 
+the user is in their program. Not only is this helpful for new programmers that need suggestions for what to write next 
 in their code, this system allows for expert programmers to write code with a great deal more ease and efficiency.
+
 \subsection{Motivation}
@@
 % overview of markov chain
+A primary motivation for undertaking this project was trying to find a new and interesting way to apply a Markov Chain.
+A Markov Chain is a stochastic process that uses the current situation and a set of probabilities to determine what 
+situation comes next. In the most simple case of a Markov Chain, the next action is based only upon the current situation
+and all other previous actions are forgotten. Thus, this process is essentially memoryless because all past states can
+be forgotten once a new state is reached. A good example of a Markov Chain is a random walk where at each integer value,
+the process chooses a new value such that the two possible values are the current value plus 1 and the current value
+minus 1. The transition probabilities for both of these options is .05. This is a good example of a Markov Chain because
+the next value on the random walk only depends on the current value in the walk; all previous values are not needed 
+to choose the next value \cite{markov}.
+
 % application of markov chain to abstact syntrax tree
@@
 
+  \begin{thebibliography}{1}
+
+  \bibitem{markov} Takis Konstantopoulos {\em Introductory lecture notes on
+Markov Chains and Random Walks} Uppsala University, URL = http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf
+
+  \end{thebibliography}
+
 \end{document}

### 2014-04-20 web/main.js
Replacement works

@@ -1,6 +1,8 @@
 var parse = require("../parse");
 var generate = require("../generate");
 var escodegen = require("escodegen");
+var esprima = require('esprima');
+
 
 var escodegenOptions = {
 	format: {
@@ -102,6 +104,7 @@ window.insertSnippet = function (node) {
 	{
 		if (isNaN(currentPosition.idx)) {
 			currentPosition.container[currentPosition.idx] = node;
+			return true;
 		} else {
 			currentPosition.container.splice(currentPosition.idx+1, 0, node);
 		}
@@ -116,11 +119,12 @@ function generateProgramSource() {
 		return;
 	}
 	var source;
-	//try {
+	try {
 		source = escodegen.generate(syntax, escodegenOptions);
-	// } catch(e) {
-	// 	console.log("generate", e);
-	// }
+	} catch(e) {
+		console.log("generate", e);
+	}
+
 	return source;
 }
 

### 2014-04-20 web/ui.js

@@ -72,22 +72,34 @@ function complete(option)
 	var node = $($(option).children()[0]).data('node');
 	console.log("snippet = ",$($(option).children()[0])," node = ",node);
 
-	window.insertSnippet(node);
+	var replaced = window.insertSnippet(node);
+
 
 	if (node.type != "Program")
 		window.setCurrentPathToNode(node);
 	
 	var snippet = $(option).children();
-	snippet.each(function () {
-		if (numspaces > 0)
-			//$(this).prepend($("<span>"+spaces+"</span>"));
-			$(this).css('padding-left','16px').css('display','inline-block');
-		$(this).insertAfter(cursor);
-	});
-	$("<br>").insertAfter(cursor);
+
+	if (replaced)
+	{
+		cursor.prev().replaceWith(snippet);
+		snippet = null;
+	}
+	else
+	{
+		snippet.each(function () {
+			if (numspaces > 0)
+				//$(this).prepend($("<span>"+spaces+"</span>"));
+				$(this).css('padding-left','16px').css('display','inline-block');
+			$(this).insertAfter(cursor);
+		});
+		$("<br>").insertAfter(cursor);
+	}
+	
 	newOptions();
 
-	cursor.insertAfter(snippet);
+	if (snippet)
+		cursor.insertAfter(snippet);
 }
 
 function moveCursorToPath(path)

### 2014-04-20 web/ui.js
Implement moving cursor

@@ -27,6 +27,16 @@ function checkKey(e)
 	        // down arrow
 	    	currentSelection += 1;
 	    }
+		else if (e.keyCode == 37)
+		{
+			// left arrow
+			moveCursor(false);
+	    }
+		else if (e.keyCode == 39)
+		{
+			// right arrow
+			moveCursor(true);
+		}
 
 	    $($('#options').children()[currentSelection]).find('pre').css('background-color','#D7EBFC');
     }
@@ -57,6 +67,53 @@ function addHandlers() {
 			e.stopPropagation();
 		}
 	});
+}
+
+// credit: https://stackoverflow.com/a/5386150
+jQuery.fn.reverse = [].reverse;
+
+jQuery.fn.eachDir = function(fn, forward) {
+	var obj = forward ? this : this.reverse();
+	return obj.each(fn);
+};
+
+function moveCursor(forward) {
+	var dirAll = forward ? 'nextAll' : 'prevAll';
+	// find the next/previous statement
+	var node, elem, found;
+	function findStatement(i, el) {
+		var n = $(el).data('node');
+		console.log("NODE!", n, el.innerText);
+		if (n) {
+			console.log("go to children");
+			$(el).children('.statement').eachDir(findStatement, forward);
+			if (found) return false;
+		}
+		if (!n) return;
+		if (isStatement(n.type))
+		{
+			node = n;
+			elem = el;
+			found = true;
+			return false;
+		}
+	}
+	var from = cursor.prev();
+	var i = 0;
+	while (!found && from.length) {
+		if (i++ > 50) throw new Error("too much recursion");
+		from[dirAll]('.statement').each(findStatement);
+		console.log("go to parent", from);
+		from = from.parent();
+	}
+	if (!from) {
+		// set it to the beginning
+	}
+	console.log("Found:", found, "elem", elem.innerText);
+	if (!node) return;
+	window.setCurrentPathToNode(node);
+	cursor.insertAfter(elem);
+	newOptions();
 }
 
 function complete(option)

### 2014-04-20 build-corpus.js
Update depth to 4

@@ -6,7 +6,7 @@ var modelDest = "web/corpus-model.js";
 var model = {};
 
 var space = +process.argv[2] || 0;
-var depth = 2;
+var depth = 4;
 
 var filesDir = "js-files";
 fs.readdirSync(filesDir).forEach(function (name) {

### 2014-04-20 web/main.js

@@ -12,6 +12,7 @@ var escodegenOptions = {
 	}
 };
 var END = "_end";
+var depth = 4;
 
 var model = window.corpusModel;
 

### 2014-04-20 generate.js
Add SequenceExpression

@@ -42,6 +42,7 @@ var nodeFeatures =
 'EmptyStatement': [],
 'ContinueStatement':[],
 'BreakStatement':[],
+'SequenceExpression':['expressions'],
 '_end':[]
 };
 
@@ -190,7 +191,8 @@ function instantiateNode(type, model, path)
 
             if (feature == "properties" || feature == "arguments" ||
                 feature == "params" || feature == "declarations" || 
-                feature == "elements" || feature == "guardedHandlers" || feature == "handlers" ||
+                feature == "elements" || feature == "guardedHandlers" ||
+				feature == "handlers" || feature == "expressions" ||
                 (feature == "body" && type == "BlockStatement")) //only BlockStatement's body is a list, Program's body is a BlockStatement
             {
                 node[feature] = generateList(model, newPath);

### 2014-04-20 parse.js

@@ -45,7 +45,8 @@ var nodeFeatures =
 'Program':['body'],
 'UpdateExpression':['operator','argument'],
 'ContinueStatement':[],
-'BreakStatement':[]
+'BreakStatement':[],
+'SequenceExpression':['expressions']
 };
 
 function traverse(path)

### 2014-04-20 web/main.js

@@ -115,7 +115,7 @@ window.insertSnippet = function (node) {
 function generateProgramSource() {
 	console.log("**** generating snippet from path ",currentPosition.path);
 
-	var syntax = generate.generateNode(model, currentPosition.path, true, 2);
+	var syntax = generate.generateNode(model, currentPosition.path, true, depth);
 	if (!syntax || syntax.type == END) {
 		return;
 	}
@@ -123,9 +123,8 @@ function generateProgramSource() {
 	try {
 		source = escodegen.generate(syntax, escodegenOptions);
 	} catch(e) {
-		console.log("generate", e);
+		console.error("generate", e);
 	}
-
 	return source;
 }
 

### 2014-04-20 web/ui.js
Add space for refreshing snippets

@@ -9,36 +9,44 @@ function checkKey(e)
 
     var selectedDiv = $($('#options').children()[currentSelection]).find('pre');
 
-    if (e.keyCode == '13') //enter key
+    if (e.keyCode == 13)
     {
+		//enter key
     	complete(selectedDiv);
     }
-    else
+	else if (e.keyCode == 32)
+	{
+		// space key
+		newOptions();
+		e.preventDefault();
+    }
+	else if (e.keyCode == 37)
+	{
+		// left arrow
+		moveCursor(false);
+	}
+	else if (e.keyCode == 39)
+	{
+		// right arrow
+		moveCursor(true);
+	}
+    else if (e.keyCode == 38 || e.keyCode == 40)
     {
 	    $($('#options').children()[currentSelection]).find('pre').css('background-color','clear');
 
-	    if (e.keyCode == '38' && currentSelection > 0)
+	    if (e.keyCode == 38 && currentSelection > 0)
 	    {
 	        // up arrow
 			currentSelection -= 1;
 	    }
-	    else if (e.keyCode == '40' && currentSelection < numSelections-1)
+	    else if (e.keyCode == 40 && currentSelection < numSelections-1)
 	    {
 	        // down arrow
 	    	currentSelection += 1;
 	    }
-		else if (e.keyCode == 37)
-		{
-			// left arrow
-			moveCursor(false);
-	    }
-		else if (e.keyCode == 39)
-		{
-			// right arrow
-			moveCursor(true);
-		}
 
 	    $($('#options').children()[currentSelection]).find('pre').css('background-color','#D7EBFC');
+		e.preventDefault();
     }
 }
 

### 2014-04-20 web/ui.js
Run button

@@ -202,4 +202,19 @@ $(document).ready(function() {
 	cursor = $('#cursor');
 	addHandlers();
 	newOptions();
+
+	$('#output').hide();
+	$(".run-btn").click(function () {
+		$('#output').show();
+		var program = $('#editor').text();
+		var ret;
+		try {
+			//var require = window.require;
+			ret = eval(program);
+		}
+		catch (e) {
+			ret = e;
+		}
+		$('#output').html(ret);
+	});
 });

### 2014-04-20 paper/paper.tex
Update paper.tex

@@
 \usepackage{float}
+\usepackage{hyperref}
 
@@
 	\bibitem{markov} Takis Konstantopoulos {\em Introductory lecture notes on
-Markov Chains and Random Walks} Uppsala University, URL = http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf
+	Markov Chains and Random Walks} Uppsala University, URL = \url{http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf}
 

### 2014-04-20 paper/paper.tex
Update paper.tex

@@
 A primary motivation for undertaking this project was trying to find a new and interesting way to apply a Markov Chain.
-A Markov Chain is a stochastic process that uses the current situation and a set of probabilities to determine what 
-situation comes next. In the most simple case of a Markov Chain, the next action is based only upon the current situation
+A Markov Chain is a stochastic process that uses the current state and a set of probabilities to determine what 
+state comes next. In the most simple case of a Markov Chain, the next action is based only upon the current situation
 and all other previous actions are forgotten. Thus, this process is essentially memoryless because all past states can
@@
 the process chooses a new value such that the two possible values are the current value plus 1 and the current value
-minus 1. The transition probabilities for both of these options is .05. This is a good example of a Markov Chain because
+minus 1. The transition probabilities for both of these options is .5. This is a good example of a Markov Chain because
 the next value on the random walk only depends on the current value in the walk; all previous values are not needed 

### 2014-04-20 paper/paper.tex
Add abstract
Reflow paper to 80 chars wide
Add more sections to the paper

@@
 
-%\begin{abstract}
-	%Your abstract goes here...
-	%...
-%\end{abstract}
+\begin{abstract}
+	We present Probabilistic Programming, a system for modeling abstract syntax
+	trees of programs with markov chains, which we use to generate new programs.
+	We attempt to use this as an assistive device to guide programmers through
+	constructing JavaScript programs.
+\end{abstract}
 
 \section{Introduction}
-For many people just starting out, programming can seem like a confusing and tedious activity. It can be unclear for 
-many new programmers where they should begin and what should come next while coding. Without any proper guidance, 
-a novice programmer is left on their own to scour thousands of pages online to discover what should come next in 
-their code. Then, once a person becomes more adept at programming, he or she may notice that they are typing out 
-the same or similar lines of code within each new program. This repetitiveness slows down the user and can make 
-programming a much more tiresome task than it needs to be. These observations steered our group towards developing
-a way for someone to code with more instruction and productivity. The Probabilistic Programming method allows for a user to code in JavaScript
-using a web application. This web app provides a list of suggestions for the user's next line of code depending on where
-the user is in their program. Not only is this helpful for new programmers that need suggestions for what to write next 
-in their code, this system allows for expert programmers to write code with a great deal more ease and efficiency.
+For many people just starting out, programming can seem like a confusing and
+tedious activity. It can be unclear for many new programmers where they should
+begin and what should come next while coding. Without any proper guidance, a
+novice programmer is left on their own to scour thousands of pages online to
+discover what should come next in their code. Then, once a person becomes more
+adept at programming, he or she may notice that they are typing out the same or
+similar lines of code within each new program. This repetitiveness slows down
+the user and can make programming a much more tiresome task than it needs to be.
+These observations steered our group towards developing a way for someone to
+code with more instruction and productivity. The Probabilistic Programming
+method allows for a user to code in JavaScript using a web application. This web
+app provides a list of suggestions for the user's next line of code depending on
+where the user is in their program. Not only is this helpful for new programmers
+that need suggestions for what to write next in their code, this system allows
+for expert programmers to write code with a great deal more ease and efficiency.
 
@@
 % overview of markov chain
-A primary motivation for undertaking this project was trying to find a new and interesting way to apply a Markov Chain.
-A Markov Chain is a stochastic process that uses the current state and a set of probabilities to determine what 
-state comes next. In the most simple case of a Markov Chain, the next action is based only upon the current situation
-and all other previous actions are forgotten. Thus, this process is essentially memoryless because all past states can
-be forgotten once a new state is reached. A good example of a Markov Chain is a random walk where at each integer value,
-the process chooses a new value such that the two possible values are the current value plus 1 and the current value
-minus 1. The transition probabilities for both of these options is .5. This is a good example of a Markov Chain because
-the next value on the random walk only depends on the current value in the walk; all previous values are not needed 
-to choose the next value \cite{markov}.
+A primary motivation for undertaking this project was trying to find a new and
+interesting way to apply a Markov Chain.  A Markov Chain is a stochastic process
+that uses the current state and a set of probabilities to determine what state
+comes next. In the most simple case of a Markov Chain, the next action is based
+only upon the current situation and all other previous actions are forgotten.
+Thus, this process is essentially memoryless because all past states can be
+forgotten once a new state is reached. A good example of a Markov Chain is a
+random walk where at each integer value, the process chooses a new value such
+that the two possible values are the current value plus 1 and the current value
+minus 1. The transition probabilities for both of these options is .5. This is a
+good example of a Markov Chain because the next value on the random walk only
+depends on the current value in the walk; all previous values are not needed to
+choose the next value \cite{markov}.
 
@@
 \section{Implementation}
+
+\subsection{Modeling}
+
+Representation of programs in a markov model
+
+Other applications of the model: genetic algorithms
+
+\subsection{Application}
+
+The vim autocomplete script
+
+The guided IDE
+
+Design decisions, HCI factors
+
+\section{Results}
+
+\subsection{Using the IDE}
+
+\subsection{Limitations of the IDE}
+
+Usability issues
+
+\subsection{Limitations of the model}
+
+Choosing valid identifiers, matching identifiers
+
+\section{Conclusions \& Open Problems}
 

### 2014-04-21 paper/paper.tex
Add some discussion of model

@@
 \usepackage{hyperref}
+\usepackage{listings}
+
+\lstset{
+	basicstyle=\footnotesize\ttfamily,
+}
 
@@
 
-Representation of programs in a markov model
+\subsubsection{AST}
+
+We work with programs at the level of abstract syntax tree (AST), using the
+syntax tree format of Mozilla's Parser API \cite{parser api}. We use
+esprima \cite{esprima}, an ECMAScript parser, to transform program source code into AST,
+and escodegen \cite{escodegen} (ECMAScript code generator) to transform ASTs into program source
+code.
+
+\subsubsection{Model}
+
+We encode the structure of an AST as a probabilistic model. The model is a data
+structure that maps node paths to probability distributions of node types or
+values. A node path is a list of node types and relationships that encodes the
+unique position of a node in a program's AST. The probability distribution for a
+node path is the set of possible node types or values that would appear in the
+syntax tree following the given path, and the respective probabilities of their
+occurance following the given path.
+
+%\lstinputlisting[caption=Scheduler, style=json]{../rand.js}
 
@@
 
+\clearpage
 \begin{thebibliography}{1}
 
+	\bibitem{esprima} Ariya Hidayat {\em Esprima}
+		URL = \url{http://esprima.org/}
+
 	\bibitem{markov} Takis Konstantopoulos {\em Introductory lecture notes on
-	Markov Chains and Random Walks} Uppsala University, URL = \url{http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf}
+		Markov Chains and Random Walks} Uppsala University,
+		URL = \url{http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf}
+
+	\bibitem{escodegen} Yusuke Suzuki {\em Escodegen}
+		URL = \url{https://github.com/Constellation/escodegen}
+
+	\bibitem{parser api} Mozilla Developer Network {\em SpiderMonkey Parser API}
+		URL = \url{https://developer.mozilla.org/en-US/docs/SpiderMonkey/Parser_API}
 
@@
 
+\clearpage
+\section*{Appendix}
+
+\begin{figure}[h!]
+	\caption{model-sample.json}
+	\centering
+	A model generated from the program ``var x = 34"
+\end{figure}
+
+\lstinputlisting{model-sample.json}
+
 \end{document}

### 2014-04-21 paper/paper.tex
Regenerate on insert

--- 2014-04-20 paper/paper.tex
@@
 that need suggestions for what to write next in their code, this system allows
-for expert programmers to write code with a great deal more ease and efficiency.
+for expert programmers to write code with more ease and efficiency.
 

### 2014-04-21 web/main.js

@@ -83,6 +83,8 @@ function setCurrentPath(path) {
 
 window.setCurrentPathToNode = function (node) {
 
+	console.log("setting current path to node ",node);
+
 	if (node.type == "ExpressionStatement")
 		node = node.expression;
 
@@ -92,10 +94,13 @@ window.setCurrentPathToNode = function (node) {
 		console.log("tree:", JSON.stringify(tree), 'Node:', JSON.stringify(node));
 	}
 
-	// console.log("current position is ",currentPosition);
+	console.log("current position is ",currentPosition);
 }
 
 window.insertSnippet = function (node) {
+
+	console.log("inserting snippet", node);
+
 	if (!tree)
 	{
 		tree = node;
@@ -103,18 +108,21 @@ window.insertSnippet = function (node) {
 	}
 	else
 	{
-		if (isNaN(currentPosition.idx)) {
+		if (isNaN(currentPosition.idx)) 
+		{
 			currentPosition.container[currentPosition.idx] = node;
-			return true;
-		} else {
+		} 
+		else 
+		{
 			currentPosition.container.splice(currentPosition.idx+1, 0, node);
 		}
 	}
+
+	//console.log("generating tree", tree);
+	return escodegen.generate(tree, escodegenOptions);
 }
 
 function generateProgramSource() {
-	console.log("**** generating snippet from path ",currentPosition.path);
-
 	var syntax = generate.generateNode(model, currentPosition.path, true, depth);
 	if (!syntax || syntax.type == END) {
 		return;
@@ -129,9 +137,28 @@ function generateProgramSource() {
 }
 
 window.getSnippets = function (num) {
+	console.log("**** generating snippet from path ",currentPosition.path);
 	return array(num).map(generateProgramSource).filter(Boolean);
 };
 
 window.getCurrentPosition = function () {
 	return currentPosition;
+}
+
+
+window.deleteSnippet = function (node) {
+
+	console.log("deleting snippet", node);
+
+	if (isNaN(currentPosition.idx)) 
+	{
+		currentPosition.container[currentPosition.idx] = null;
+	} 
+	else
+	{
+		currentPosition.container.splice(currentPosition.idx, currentPosition.idx);
+	}
+
+	//console.log("generating tree", tree);
+	return escodegen.generate(tree, escodegenOptions);
 }

### 2014-04-21 web/ui.js

@@ -32,6 +32,13 @@ function checkKey(e)
 		// right arrow
 		moveCursor(true);
 	}
+	else if (e.keyCode == 8)
+	{
+		e.preventDefault();
+
+		// delete key
+		deleteKey();
+	}
     else if (e.keyCode == 38 || e.keyCode == 40)
     {
 	    $($('#options').children()[currentSelection]).find('pre').css('background-color','clear');
@@ -118,7 +125,7 @@ function moveCursor(forward) {
 	if (!from) {
 		// set it to the beginning
 	}
-	console.log("Found:", found, "elem", elem.innerText);
+	console.log("Found:", found, "elem", elem && elem.innerText);
 	if (!node) return;
 	window.setCurrentPathToNode(node);
 	cursor.insertAfter(elem);
@@ -133,58 +140,61 @@ function scrollIntoView(container, elem) {
 	}
 }
 
-function complete(option)
-{
-	var selectedLine = cursor.parent().text();
-	selectedLine = selectedLine.replace(/^\{\n/, "");
-	var spaces = "";
-	var numspaces = selectedLine.match(/^ */)[0].length;
-	if (numspaces == 1) numspaces = 0;
-	for (var i = 0; i < numspaces; i++)
-		spaces += " ";
-	console.log("line = ",selectedLine,"  spaces = ",numspaces);
-
-
-	// console.log("generating from position ",window.getCurrentPosition());
-
-	// console.log("cursor");
-
-	var node = $($(option).children()[0]).data('node');
-	console.log("snippet = ",$($(option).children()[0])," node = ",node);
-
-	var replaced = window.insertSnippet(node);
-
-
-	if (node.type != "Program")
-		window.setCurrentPathToNode(node);
-	
-	var snippet = $(option).children();
-
-	if (replaced)
-	{
-		cursor.prev().replaceWith(snippet);
-		snippet = null;
-	}
-	else
-	{
-		snippet.each(function () {
-			if (numspaces > 0)
-				//$(this).prepend($("<span>"+spaces+"</span>"));
-				$(this).css('padding-left','16px').css('display','inline-block');
-			$(this).insertAfter(cursor);
-		});
-		$("<br>").insertAfter(cursor);
-	}
-	
-	newOptions();
-
-	if (snippet)
-		cursor.insertAfter(snippet);
+function objectsEqual(a, b) {
+	return JSON.stringify(a) == JSON.stringify(b);
 }
 
-function moveCursorToPath(path)
[…]

### 2014-04-21 web/main.js
Look up nodes from DOM by ID instead of JSON

@@ -32,6 +32,20 @@ function objectsEqual(a, b) {
 	return JSON.stringify(a) == JSON.stringify(b);
 }
 
+function walkSyntax(syntax, fn) {
+	if (!syntax || typeof syntax != "object") {
+		return;
+	}
+	if (syntax.type) {
+		fn(syntax);
+	}
+	for (var key in syntax) {
+		if (key != "type" && hasOwnProp(syntax, key)) {
+			walkSyntax(syntax[key], fn);
+		}
+	}
+}
+
 var FOLLOW = "F";
 // looking for node in syntax
 function getPath(syntax, node, depth, path, container, idx) {
@@ -122,11 +136,27 @@ window.insertSnippet = function (node) {
 	return escodegen.generate(tree, escodegenOptions);
 }
 
+var maxId = 0;
+var nodesById = {};
+function labelNodeId(node) {
+	node.idX = ++maxId;
+	nodesById[node.idX] = node;
+}
+
+window.getNodeById = function (id) {
+	var node = nodesById[id];
+	if (!node) {
+		console.log("no node found with id", id);
+	}
+	return node;
+};
+
 function generateProgramSource() {
 	var syntax = generate.generateNode(model, currentPosition.path, true, depth);
 	if (!syntax || syntax.type == END) {
 		return;
 	}
+	walkSyntax(syntax, labelNodeId);
 	var source;
 	try {
 		source = escodegen.generate(syntax, escodegenOptions);
@@ -143,8 +173,7 @@ window.getSnippets = function (num) {
 
 window.getCurrentPosition = function () {
 	return currentPosition;
-}
-
+};
 
 window.deleteSnippet = function (node) {
 

### 2014-04-21 web/ui.js

@@ -74,7 +74,7 @@ function addHandlers() {
 	});
 	editor.on('click', '.statement', function(e) {
 		var el = $(this);
-		var node = el.data('node');
+		var node = window.getNodeById(el.data('node-id'));
 		if (isStatement(node.type))
 		{
 			window.setCurrentPathToNode(node);
@@ -98,7 +98,7 @@ function moveCursor(forward) {
 	// find the next/previous statement
 	var node, elem, found;
 	function findStatement(i, el) {
-		var n = $(el).data('node');
+		var n = window.getNodeById($(el).data('node-id'));
 		console.log("NODE!", n, el.innerText);
 		if (n) {
 			console.log("go to children");
@@ -148,7 +148,7 @@ function findNode(span, node)
 {
 	span = $(span);
 
-	if (objectsEqual(span.data('node'), node))
+	if (window.getNodeById(span.data('node-id')) == node)
 		return span;
 
 	var children = span.children();
@@ -162,7 +162,7 @@ function findNode(span, node)
 
 function complete(option)
 {
-	var node = $($(option).children()[0]).data('node');
+	var node = window.getNodeById($($(option).children()[0]).data('node-id'));
 	var newCode = window.insertSnippet(node);
 
 	$('#editor').html(newCode);

### 2014-04-21 escodegen-snippet.js
Add esprima patch. Compare nodes directly.

@@ -1,0 +1,3 @@
+// line 2026
+        var nodeId = stmt.idX;
+        result = ["<span class='statement' data-node-id='"+nodeId+"'>"].concat(result, "</span>");

### 2014-04-21 web/main.js

@@ -28,10 +28,6 @@ function array(length) {
 
 var hasOwnProp = Function.prototype.call.bind(Object.prototype.hasOwnProperty);
 
-function objectsEqual(a, b) {
-	return JSON.stringify(a) == JSON.stringify(b);
-}
-
 function walkSyntax(syntax, fn) {
 	if (!syntax || typeof syntax != "object") {
 		return;
@@ -58,7 +54,7 @@ function getPath(syntax, node, depth, path, container, idx) {
 	if (syntax.type == "ExpressionStatement")
 		syntax = syntax.expression;
 
-	if (objectsEqual(syntax, node)) {
+	if (syntax == node) {
 		if (path[path.length-1] == FOLLOW || path[path.length-1] == 'body')
 			path = path.concat(node.type, FOLLOW)
 		return {"path":path, "container":container, "idx":idx};

### 2014-04-21 web/main.js
Don't allow adding no snippet

@@ -148,13 +148,14 @@ window.getNodeById = function (id) {
 };
 
 function generateProgramSource() {
-	var syntax = generate.generateNode(model, currentPosition.path, true, depth);
-	if (!syntax || syntax.type == END) {
-		return;
-	}
-	walkSyntax(syntax, labelNodeId);
 	var source;
 	try {
+		var syntax = generate.generateNode(model, currentPosition.path, true, depth);
+		if (!syntax || syntax.type == END) {
+			return;
+		}
+		walkSyntax(syntax, labelNodeId);
+		
 		source = escodegen.generate(syntax, escodegenOptions);
 	} catch(e) {
 		console.error("generate", e);

### 2014-04-21 web/ui.js

@@ -14,7 +14,8 @@ function checkKey(e)
     if (e.keyCode == 13)
     {
 		//enter key
-    	complete(selectedDiv);
+		if (numSelections > 0)
+	    	complete(selectedDiv);
     }
 	else if (e.keyCode == 32)
 	{
@@ -48,7 +49,7 @@ function checkKey(e)
 	        // up arrow
 			currentSelection -= 1;
 	    }
-	    else if (e.keyCode == 40 && currentSelection < numSelections-1)
+	    else if (e.keyCode == 40 && currentSelection < numSelections)
 	    {
 	        // down arrow
 	    	currentSelection += 1;

### 2014-04-21 paper/paper.tex
Add screenshots and discussion of vim autocomplete

@@
 \usepackage{graphicx}
+\usepackage{subcaption}
 \usepackage{float}
@@
 syntax tree format of Mozilla's Parser API \cite{parser api}. We use
-esprima \cite{esprima}, an ECMAScript parser, to transform program source code into AST,
+esprima \cite{esprima}, an ECMAScript parser, to transform program source code into an AST,
 and escodegen \cite{escodegen} (ECMAScript code generator) to transform ASTs into program source
@@
 syntax tree following the given path, and the respective probabilities of their
-occurance following the given path.
+occurrence following the given path.
+
+The model is implemented as a JSON object where the path is indexed into the
+object by successively indexing by each string of the path.
+
+See figure~\ref{fig:sample-model} in the appendix for a model for the program
+``var x = 34;".
 
@@
 
-The vim autocomplete script
+\subsubsection{vim autocomplete script}
 
-The guided IDE
+We integrated our model into the text editor vim to offer autocompletion
+suggestions of programs upon the user's request, using vim's
+\texttt{completefunc} feature. When the user is editing a JS file in vim with
+this feature, they may press a keyboard command to get a list of program
+snippets to choose from. The snippets are generated in the context of the
+program path at cursor's position. The user may pick one of the snippet
+suggestions to have it inserted into their code at their cursor.
+
+The autocomplete snippets are generated as follows:
+\begin{enumerate}
+	\item Parse the contents of the current file into an AST.
+	\item Find the node path corresponding to the cursor by traversing the AST.
+		%This represents the program context at which the code snippets should
+		%be generated.
+	\item Generate programs starting from the node path of the cursor.
+	\item Present the programs to the user as code completion snippets.
+\end{enumerate}
+
+The vim autocomplete feature is limited to generating full programs on a single
+line, and requires explicitly triggering the \texttt{completefunc} feature for
+code snippets.
+
+\begin{figure}[h]
+	\centering
+	\begin{tabular}{c c}
+
+		\begin{subfigure}[h]{6.25cm}
+			\centering
+			\includegraphics[width=1.00\textwidth]{vim1}
+			\caption{Inserting a for loop} \label{fig:vim1}
+		\end{subfigure}
+
+		\begin{subfigure}[h]{6cm}
+			\centering
+			\includegraphics[width=1.00\textwidth]{vim2}
+			\caption{Inserting an object property} \label{fig:vim2}
+		\end{subfigure}
+	\end{tabular}
+
+	\caption{Screenshots of vim using autocomplete}
+\end{figure}
+
+\subsubsection{Guided IDE}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.90\textwidth]{screenshot}
+  \caption{Screenshot of the web program editor GUI} \label{fig:screenshot}
+\end{figure}
 
@@
 	\caption{model-sample.json}
+	\label{fig:sample-model}
 	\centering
-	A model generated from the program ``var x = 34"
+	A model generated from the program ``var x = 34;"
 \end{figure}

### 2014-04-21 paper/paper.tex
Update paper.tex

--- 2014-04-21 paper/paper.tex
@@
 syntax tree following the given path, and the respective probabilities of their
-occurance following the given path.
+occurrence following the given path.
 

### 2014-04-21 paper/paper.tex
Further discuss vim autocomplete and web interface

--- 2014-04-21 paper/paper.tex
@@
 We work with programs at the level of abstract syntax tree (AST), using the
-syntax tree format of Mozilla's Parser API \cite{parser api}. We use
-esprima \cite{esprima}, an ECMAScript parser, to transform program source code into an AST,
-and escodegen \cite{escodegen} (ECMAScript code generator) to transform ASTs into program source
-code.
+syntax tree format of Mozilla's Parser API \cite{parser api}. We use esprima
+\cite{esprima}, an ECMAScript parser, to transform program source code into an
+AST, and escodegen \cite{escodegen} (ECMAScript code generator) to transform
+ASTs into program source code.
 
@@
 syntax tree following the given path, and the respective probabilities of their
-occurrence following the given path.
+occurance following the given path.
 
@@
 See figure~\ref{fig:sample-model} in the appendix for a model for the program
-``var x = 34;".
+``var x = 34;" In this model we see for example that the path
+\texttt{[Program, body, VariableDeclaration, declarations]} leads to
+\texttt{VariableDeclarator} with probability 1. This indicates that a Program
+node whose body is a VariableDeclaration always has a declaration that is of
+type VariableDeclarator.
 
-%\lstinputlisting[caption=Scheduler, style=json]{../rand.js}
-
-Other applications of the model: genetic algorithms
-
+\clearpage
 \subsection{Application}
@@
 
-The vim autocomplete feature is limited to generating full programs on a single
-line, and requires explicitly triggering the \texttt{completefunc} feature for
-code snippets.
-
 \begin{figure}[h]
@@
 	\begin{tabular}{c c}
-
 		\begin{subfigure}[h]{6.25cm}
@@
 		\end{subfigure}
-
 		\begin{subfigure}[h]{6cm}
@@
 	\end{tabular}
-
-	\caption{Screenshots of vim using autocomplete}
+	\caption{Screenshots of vim using autocomplete of programs}
 \end{figure}
 
+The vim autocomplete feature is limited to inserting program snippets on a single
+line, and requires the user to explicitly trigger the \texttt{completefunc}
+feature to be prompted with code snippets. To further explore using our model
+for editing, we created a web-based editor that more deeply integrates Probabilistic
+Programming.
+
+\clearpage
 \subsubsection{Guided IDE}
+
+The Guided IDE offers users an environment for building a program entirely using
+our model, using program snippets generated at the cursor's context as building
+blocks. The editor presents a window for the program text, and and list of
+suggestion code snippets. The user can move a cursor through the code window at
+positions corresponding to boundaries of program nodes. As the user moves the
+cursor, the snippet suggestions are regenerated using the cursor's node path, to
+continuosly provide suggestions relevant to the current context of the program. When the
+user selects a suggestion snippet, its AST is inserted into the program's AST at
+the cursor's position in the tree, and the code window is redrawn. The user can
+move the cursor within newly inserted blocks, so that the program can grow
+deeper. The initial corpus for the IDE's model consists of the JavaScript files
+that comprise our project.
 
@@
 
-Design decisions, HCI factors
+%Design decisions, HCI factors
 
@@
 \subsection{Using the IDE}
+
+We can use the Guided IDE to generate programs.
 
@@
 \section{Conclusions \& Open Problems}
+
+Other applications of the model: genetic algorithms
 

### 2014-04-21 paper/paper.tex
Discuss Guided IDE more

@@
 \usepackage{listings}
+
+\hyphenation{ECMAScript JavaScript escodegen esprima}
 
@@
 
-We can use the Guided IDE to generate programs.
+We can use the Guided IDE to generate programs. The programs look like actual
+programs but do not actually function unless the user is very careful about
+which snippets they pick. Crucially, the variable and identifier names that our
+system generates are not in general meaningful in the scope of the generated
+program.  Whereas with the vim autocompletions the identifiers would often be
+valid because because the corpus was the current file, with the larger corpus of
+the Guided IDE and writing a program from scratch, the user is less likely to
+find program snippets that use valid identifiers.
+
+This is a limitation of the current model, that identifier names are chosen
+probabilistically based on their position in the code, without regard to whether
+they are in scope and without refering to anything in another part of the
+program.
 

### 2014-04-21 paper/paper.tex
Discuss alternative IDE approaches

@@
   \centering
-  \includegraphics[width=0.90\textwidth]{screenshot}
+  \includegraphics[width=1.00\textwidth]{screenshot}
   \caption{Screenshot of the web program editor GUI} \label{fig:screenshot}
@@
 they are in scope and without refering to anything in another part of the
-program.
+program. Because of this limitation, the Guided IDE is not useful for practical
+purposes. We consider some alternative approaches for the Guided IDE using a
+similar model.
 
-\subsection{Limitations of the IDE}
+\subsection{Alternative approaches for the IDE}
 
-Usability issues
+\subsubsection{Prompt for identifiers}
+
+Rather than picking identifier names from the model, use placeholder identifiers
+and prompt the user to enter actual identifiers, with suggestions based on the
+current variables in scope and previously used identifiers. This would give the
+user the opportunity to create semantically valid programs, but the entering of
+identifier names could be tedious.
+
+\subsubsection{Stubs}
+
+Instead of having a movable cursor in the code window, have stubs in the
+program. Each stub would be a marker in place of an AST node. The user would
+select a stub and be given choices of snippets to replace it. Selecting a
+snippet would move the current snippet into the program in place of the selected
+stub.  The generated snippets could themselves have stubs in them, so that the
+user may continue to complete stubs and expand their program.  At the level of
+variable identifiers, the snippet suggestions would be taken from the computed
+list of in-scope variables at the position of the selected stub, unless the
+completion was for a variable declaration, in which case the user would be
+prompted to name a new variable.
+
+This approach, while offering less places to insert code then the current
+approach, would allow for programs to more closely be generated from our model,
+since the stub completions would always be at leaf nodes of the AST, whereas in
+the current approach the user may choose to insert a snippet in the middle of a
+list, which means that the transition from the inserted snippet to the following
+node in the list does not in general follow from the model.
+
+While the stub approach would be more restrictive for the user as it requires
+them to build the program in a certain way, with proper handling of identifiers,
+it may provide for an interesting or useful paradigm for interacting with a program
+source.
+
+%\subsection{Limitations of the IDE}
 
@@
 
-Choosing valid identifiers, matching identifiers
+\subsection{Extending the model}
+
+%Choosing valid identifiers, matching identifiers
 

### 2014-04-21 paper/paper.tex
Section 2 stuff

--- 2014-04-21 paper/paper.tex
@@
 \usepackage{amsmath}   
+\usepackage{mathtools}
 \usepackage{graphicx}
-\usepackage{subcaption}
 \usepackage{float}
@@
 \usepackage{listings}
+
+\usepackage{qtree}
+\usepackage{tikz}
+\usepackage{tikz-qtree}
+
+\usepackage[]{algorithm2e}
 
@@
 	We present Probabilistic Programming, a system for modeling abstract syntax
-	trees of programs with markov chains, which we use to generate new programs.
+	trees of programs with Markov chains, which we use to generate new programs.
 	We attempt to use this as an assistive device to guide programmers through
@@
 \section{Introduction}
+
+\subsection{Motivation}
+
 For many people just starting out, programming can seem like a confusing and
@@
 
-\subsection{Motivation}
+\subsection{Markov chain}
 
-% overview of markov chain
-A primary motivation for undertaking this project was trying to find a new and
-interesting way to apply a Markov Chain.  A Markov Chain is a stochastic process
+We chose to use a Markov chain to drive the code snippet generation, because
+(a) , and (b). A Markov Chain is a stochastic process
 that uses the current state and a set of probabilities to determine what state
@@
 
-% application of markov chain to abstact syntrax tree
+% overview of markov chain
 
-\section{Implementation}
+\section{Tree-based Markov chain}
 
-\subsection{Modeling}
+\subsection{Abstract syntax trees}
 
-\subsubsection{AST}
+An abstract syntax tree (AST) is a tree-based, rather than text-based, representation of a program.
+Such a tree is generated using the grammar rules of a programming language and text input. In the tree, \emph{nodes}
+are language control structures such as {\tt for} loops, {\tt if} statements, variable and function
+declarations, and so on. The \emph{edges} connecting a parent node to its children represent the different \emph{features}
+of that node. For instance, a {\tt for} loop would have the features \emph{init} (e.g. {\tt var i = 0;}),
+\emph{test} (e.g. {\tt i < array.length;}), \emph{update} (e.g. {\tt i++}), and a \emph{body} (a list
+of statements that should be executed per loop iteration). To implement lists, we include the additional
+\emph{follow} edge as a feature to all statements, which would connect a statement node to the next statement
+in the list. If the statement is last in its container list, we connect its \emph{follow} edge to a special
+{\tt {\_end}} node.
 
-We work with programs at the level of abstract syntax tree (AST), using the
-syntax tree format of Mozilla's Parser API \cite{parser api}. We use esprima
-\cite{esprima}, an ECMAScript parser, to transform program source code into an
-AST, and escodegen \cite{escodegen} (ECMAScript code generator) to transform
-ASTs into program source code.
+\begin{center}
 
-\subsubsection{Model}
+[example of AST here]
 
-We encode the structure of an AST as a probabilistic model. The model is a data
-structure that maps node paths to probability distributions of node types or
-values. A node path is a list of node types and relationships that encodes the
-unique position of a node in a program's AST. The probability distribution for a
-node path is the set of possible node types or values that would appear in the
+% \begin{tikzpicture}[level distance=1.5cm,
+% level 1/.style={sibling distance=3.5cm},
+% level 2/.style={sibling distance=1cm}]
+
+% \node (Root) {Program}
+% child {
+%     node {for} edge from parent node[draw=none] {\emph{body}}
+%     child { node {VD} edge from parent node[draw=none] {\emph{init}} }
+% };
+
+% \end{tikzpicture}
+
+\end{center}
+
+Our system deals with programs at the level of AST's as they provide more structure and modularity than text.
+We use the syntax tree format of Mozilla's Parser API \cite{parser api}, and \emph{esprima} \cite{esprima}, an ECMAScript
+parser, to transform program source code into an AST of this format.
+
+%escodegen \cite{escodegen} (ECMAScript code generator) to transform ASTs into program source code.
+
+\subsection{Our Markov chain}
+
+To build our Markov chain structure, we parse a corpus of JavaScript files into their
+respective AST's, and then collect statistics from these trees to build our Markov model.
+
+The model maps AST paths to the probability distributions of node types or values for that given
+path. An AST path is a list of items alternating between a node type and one of its features, thus encoding the unique position of a 
+specific node in the AST. The probability distribution for an AST path is the set of possible node types or values that would appear in the
 syntax tree following the given path, and the respective probabilities of their
@@
 
-The model is implemented as a JSON object where the path is indexed into the
-object by successively indexing by each string of the path.
+More formally, given a depth $d$ (indicating how many nodes we should backtrace) for a given
+${node}_d$ and feature ${edge}_d$, the generation of ${node}_d$'s child node given by the
+transition ${edge}_d$, can be expressed by the following function $f$,
 
-See figure~\ref{fig:sample-model} in the appendix for a model for the program
-``var x = 34;" In this model we see for example that the path
-\texttt{[Program, body, VariableDeclaration, declarations]} leads to
-\texttt{VariableDeclarator} with probability 1. This indicates that a Program
-node whose body is a VariableDeclaration always has a declaration that is of
-type VariableDeclarator.
+\begin{align*}
+f({node}_1, {edge}_1, ..., {node}_{d}, {edge}_{d}) = \begin{dcases*}
+		{newnode_1},	&	$P({newnode_1})$ \\
+		{newnode_2},	&	$P({newnode_2})$ \\
+		...\\
+		{newnode_n},    &   $P({newnode_n})$ \\
+	\end{dcases*}
+\end{align*}
 
-\clearpage
-\subsection{Application}
+such that ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., and ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_k$ given the probabilities $P({newnode_1}), ..., P({newnode_n})$, and
 
-\subsubsection{vim autocomplete script}
+\begin{align*}
+\sum\limits_{i = 0}^n P({newnode_i}) = 1.
+\end{align*}
 
-We integrated our model into the text editor vim to offer autocompletion
-suggestions of programs upon the user's request, using vim's
-\texttt{completefunc} feature. When the user is editing a JS file in vim with
-this feature, they may press a keyboard command to get a list of program
-snippets to choose from. The snippets are generated in the context of the
-program path at cursor's position. The user may pick one of the snippet
-suggestions to have it inserted into their code at their cursor.
+From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with, $\forall {edge} \in {Features}({newnode}_k)$,
 
-The autocomplete snippets are generated as follows:
-\begin{enumerate}
-	\item Parse the contents of the current file into an AST.
-	\item Find the node path corresponding to the cursor by traversing the AST.
-		%This represents the program context at which the code snippets should
-		%be generated.
-	\item Generate programs starting from the node path of the cursor.
-	\item Present the programs to the user as code completion snippets.
-\end{enumerate}
+\begin{align*}
+{Path}({newnode}_k, {edge}) = f({node}_2, {edge}_2, ..., {node}_{d}, {edge}_{d}, {newnode}_k, {edge}).
+\end{align*}
 
-\begin{figure}[h]
-	\centering
-	\begin{tabular}{c c}
-		\begin{subfigure}[h]{6.25cm}
-			\centering
-			\includegraphics[width=1.00\textwidth]{vim1}
-			\caption{Inserting a for loop} \label{fig:vim1}
-		\end{subfigure}
-		\begin{subfigure}[h]{6cm}
-			\centering
-			\includegraphics[width=1.00\textwidth]{vim2}
-			\caption{Inserting an object property} \label{fig:vim2}
-		\end{subfigure}
-	\end{tabular}
-	\caption{Screenshots of vim using autocomplete of programs}
-\end{figure}
+This process will eventually terminate as the probability of generating {\tt \_end} nodes for a \emph{follow} edge will always be non-zero -- as the programs we parse into our corpus are surely finite -- and such nodes have no features to be ``expanded.''
 
-The vim autocomplete feature is limited to inserting program snippets on a single
-line, and requires the user to explicitly trigger the \texttt{completefunc}
-feature to be prompted with code snippets. To further explore using our model
-for editing, we created a web-based editor that more deeply integrates Probabilistic
-Programming.
+\subsection{Implementation}
 
-\clearpage
-\subsubsection{Guided IDE}
+\subsubsection{Data structure}
 
-The Guided IDE offers users an environment for building a program entirely using
-our model, using program snippets generated at the cursor's context as building
-blocks. The editor presents a window for the program text, and and list of
-suggestion code snippets. The user can move a cursor through the code window at
-positions corresponding to boundaries of program nodes. As the user moves the
-cursor, the snippet suggestions are regenerated using the cursor's node path, to
-continuosly provide suggestions relevant to the current context of the program. When the
-user selects a suggestion snippet, its AST is inserted into the program's AST at
-the cursor's position in the tree, and the code window is redrawn. The user can
-move the cursor within newly inserted blocks, so that the program can grow
-deeper. The initial corpus for the IDE's model consists of the JavaScript files
-that comprise our project.
+Building the Markov chain model means being able to provide the probability distribution of new nodes for arbitrary input,
+as described by the function in Section 2.2. To do this, we constructed a data structure using JavaScript Object Notation (JSON),
+a structural format useful for ease of specification and integration with JavaScript programs.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.90\textwidth]{screenshot}
-  \caption{Screenshot of the web program editor GUI} \label{fig:screenshot}
-\end{figure}
+The data structure is a hash table (also commonly called a map, dictionary, or associative array), that, on a given input, returns another hash table for paths that begin with that input.
+This behavior is repeated until the number of inputs reaches the desired depth, at which point a final hash table, mapping the node type to its probability distribution is returned.
+This, in effect, is modularizing the function $f$ described above, so that, say for ${depth} = 2$,
 
-%Design decisions, HCI factors
+\begin{align*}
+lookup(lookup(lookup(lookup({hash}, {node}_1), {edge}_1), {node}_2), {edge}_2) = \textbf{P}({nodes}).
+\end{align*}
+
+\subsubsection{Building the model}
+
+We first obtain collection of JavaScript programs in AST form. We then run the following algorithm to construct our model:
+
+\begin{verbatim}
+procedure addToModel(node, path):
+    model = corpusModel
+    for each elem in path[path.length-depth*2 .. path.length]:
+        model = model[elem]
+    model[node.type] += 1
+    model[_total] += 1
+    for each feature of node:
+        temp_path = path + [node.type, feature]
+        if node.feature is list of nodes:
+            for each item in node.feature:
+                addToModel(item, temp_path)
+                temp_path += [item.type, "FOLLOW"]
+            addToModel(END_NODE, temp_path)
+        else:
+            addToModel(node.feature, temp_path)
+
+for each parsed AST:
+    addToModel(AST.root, [])
+\end{verbatim}
+
+For each AST, we walk through every node in the tree by recursively exploring each feature\footnote{While we use \emph{follow} as if it were a feature edge, programs encoded in the AST's we parse use arrays of nodes (rather than linked-lists as we represent them) for a collection of statements.
+The given pseudocode demonstrates how we can transform an array into a linear sequence of edges.} of each node.
+In each call to explore a node's feature, we maintain the path up until that node by appending to it the node type and the name of the feature to reach the wanted child.
+We trim the beginning of the path off so that the length of the path is appropriate with respect to the chosen depth (i.e.
+so that $|path| = 2*d$)\footnote{Note that this requires that the path is at least of length $2*depth$.
+If the path is shorter than this, it represents not having at least $d$ levels of context.
+This can occur, say, for the first statement in the entire program, where there are no preceding nodes in its AST path.
+For these cases, we add the ``{\tt \_null}'' feature until we reach the desired depth (omitted in the pseudocode.)}, create that path through the layers of hash tables, if it doesn't already exist, add the node type to the probability distribution if it isn't already, and then increment a counter associated with that node type.
+We also increment a ``{\tt \_total}'' counter for the convenience of the generator program.
+
+When this process is complete, the JSON object is saved to a file and can be augmented by more code samples thereafter. (An example of such a model is shown in the Appendix.)
+
+\subsection{Generating code}
+
+\subsubsection{Random programs}
+
+\subsubsection{Snippets}
+
+% application of markov chain to abstact syntrax tree
+
+%\lstinputlisting[caption=Scheduler, style=json]{../rand.js}
+
+Other applications of the model: genetic algorithms
+
+\section{Applications}
+
+The vim autocomplete script
+
+The guided IDE
+
+Design decisions, HCI factors
 
@@
 \subsection{Using the IDE}
-
-We can use the Guided IDE to generate programs.
 
@@
 \section{Conclusions \& Open Problems}
-
-Other applications of the model: genetic algorithms
 
@@
 	\caption{model-sample.json}
-	\label{fig:sample-model}
 	\centering
-	A model generated from the program ``var x = 34;"
+	A model generated from the program ``var x = 34"
 \end{figure}

### 2014-04-21 paper/paper.tex
Discuss high-order model in conclusions

--- 2014-04-21 paper/paper.tex
@@
 	\item Find the node path corresponding to the cursor by traversing the AST.
-		%This represents the program context at which the code snippets should
-		%be generated.
 	\item Generate programs starting from the node path of the cursor.
@@
 
-%Design decisions, HCI factors
-
 \section{Results}
@@
 
-This is a limitation of the current model, that identifier names are chosen
+\subsection{Limitations of the model}
+
+It is a limitation of the current model that identifier names are chosen
 probabilistically based on their position in the code, without regard to whether
@@
 
-%\subsection{Limitations of the IDE}
+\section{Conclusions \& Open Problems}
 
-\subsection{Limitations of the model}
+%Other applications of the model: genetic algorithms
 
@@
 
-%Choosing valid identifiers, matching identifiers
+Just as our approach of modeling a program with a tree-based Markov chain is a
+higher order model than a linear Markov chain of the program source code tokens,
+there may be applications of probabilistic modeling to higher-order
+representations of programs than the AST, such as a higher-order abstract syntax
+\cite{hosa}. Such an application may model the relationship between different
+parts of an AST, and handle scoping of variables to model identifiers
+abstractly.
 
-\section{Conclusions \& Open Problems}
-
-Other applications of the model: genetic algorithms
+With a more sophisticated model of identifiers that could generate programs with
+identifiers in meaningful relation to eachother, Probabilistic Programming could
+be used to generate semantically valid programs, for greater assistive,
+expressive, and programatic power.
 
@@
 	\bibitem{esprima} Ariya Hidayat {\em Esprima}
-		URL = \url{http://esprima.org/}
+		\\\url{http://esprima.org/}
 
 	\bibitem{markov} Takis Konstantopoulos {\em Introductory lecture notes on
-		Markov Chains and Random Walks} Uppsala University,
-		URL = \url{http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf}
+		Markov Chains and Random Walks.} Uppsala University,
+		\\\url{http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf}
+
+	\bibitem{hosa} Frank Pfenning, Conal Elliott
+		{\em Higher-Order Abstract Syntax.}
+		ACM SIGPLAN Notices. Vol. 23. No. 7. ACM, 1988.
+		\\\url{http://www.cs.cmu.edu/~fp/papers/pldi88.pdf}
 
 	\bibitem{escodegen} Yusuke Suzuki {\em Escodegen}
-		URL = \url{https://github.com/Constellation/escodegen}
+		\\\url{https://github.com/Constellation/escodegen}
 
 	\bibitem{parser api} Mozilla Developer Network {\em SpiderMonkey Parser API}
-		URL = \url{https://developer.mozilla.org/en-US/docs/SpiderMonkey/Parser_API}
+		\\\url{https://developer.mozilla.org/en-US/docs/SpiderMonkey/Parser_API}
 

### 2014-04-21 paper/paper.tex
Start generate section

--- 2014-04-21 paper/paper.tex
@@
 ${node}_d$ and feature ${edge}_d$, the generation of ${node}_d$'s child node given by the
-transition ${edge}_d$, can be expressed by the following function $f$,
-
-\begin{align*}
+transition ${edge}_d$, can be expressed by the following function $f$, \begin{align*}
 f({node}_1, {edge}_1, ..., {node}_{d}, {edge}_{d}) = \begin{dcases*}
@@
 
-such that ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., and ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_k$ given the probabilities $P({newnode_1}), ..., P({newnode_n})$, and
-
-\begin{align*}
+such that ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., and ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_k$ given the probabilities $P({newnode_1}), ..., P({newnode_n})$, and \begin{align*}
 \sum\limits_{i = 0}^n P({newnode_i}) = 1.
@@
 
-From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with, $\forall {edge} \in {Features}({newnode}_k)$,
-
-\begin{align*}
+From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with, $\forall {edge} \in {Features}({newnode}_k)$, \begin{align*}
 {Path}({newnode}_k, {edge}) = f({node}_2, {edge}_2, ..., {node}_{d}, {edge}_{d}, {newnode}_k, {edge}).
@@
 This behavior is repeated until the number of inputs reaches the desired depth, at which point a final hash table, mapping the node type to its probability distribution is returned.
-This, in effect, is modularizing the function $f$ described above, so that, say for ${depth} = 2$,
-
+This, in effect, is modularizing the function $f$ described above, so that we can reach the probability distribution with minimal code and data duplication.
+This is demonstrated by following example, for  ${depth} = 2$.
 \begin{align*}
-lookup(lookup(lookup(lookup({hash}, {node}_1), {edge}_1), {node}_2), {edge}_2) = \textbf{P}({nodes}).
+lookup(lookup(lookup(lookup({model}, {node}_1), {edge}_1), {node}_2), {edge}_2) = \textbf{P}({nodes}).
 \end{align*}
@@
 We trim the beginning of the path off so that the length of the path is appropriate with respect to the chosen depth (i.e.
-so that $|path| = 2*d$)\footnote{Note that this requires that the path is at least of length $2*depth$.
+so that $|path| = 2*d$).\footnote{Note that this requires that the path is at least of length $2*depth$.
 If the path is shorter than this, it represents not having at least $d$ levels of context.
 This can occur, say, for the first statement in the entire program, where there are no preceding nodes in its AST path.
-For these cases, we add the ``{\tt \_null}'' feature until we reach the desired depth (omitted in the pseudocode.)}, create that path through the layers of hash tables, if it doesn't already exist, add the node type to the probability distribution if it isn't already, and then increment a counter associated with that node type.
+For these cases, we add the ``{\tt \_null}'' feature until we reach the desired depth (omitted in the pseudocode.)} We traverse through the layers of hash tables, creating the path elements if they don't already exist, add the node type to the probability distribution if it isn't already, and then increment a counter associated with that node type.
 We also increment a ``{\tt \_total}'' counter for the convenience of the generator program.
@@
 \subsubsection{Random programs}
+
+Random programs can be generated rather easily using the approach explained in Section 2.2.
+The program tree is built up one node at a time, each one generated based on its unique AST path.
+We always begin a program with a ``Program'' node
 

### 2014-04-21 paper/paper.tex

@@
 \usepackage[]{algorithm2e}
+\hyphenation{ECMAScript JavaScript escodegen esprima}
 
@@
 
+See figure~\ref{fig:sample-model} in the appendix for a model for the program
+``var x = 34;" In this model we see for example that the path
+\texttt{[Program, body, VariableDeclaration, declarations]} leads to
+\texttt{VariableDeclarator} with probability 1. This indicates that a Program
+node whose body is a VariableDeclaration always has a declaration that is of
+type VariableDeclarator.
+
+The model is implemented as a JSON object where the path is indexed into the
+object by successively indexing by each string of the path.
+
 \subsection{Generating code}
@@
 
-Other applications of the model: genetic algorithms
-
 \section{Applications}
 
-The vim autocomplete script
+\subsection{vim autocomplete script}
 
-The guided IDE
+We integrated our model into the text editor vim to offer autocompletion
+suggestions of programs upon the user's request, using vim's
+\texttt{completefunc} feature. When the user is editing a JS file in vim with
+this feature, they may press a keyboard command to get a list of program
+snippets to choose from. The snippets are generated in the context of the
+program path at cursor's position. The user may pick one of the snippet
+suggestions to have it inserted into their code at their cursor.
 
-Design decisions, HCI factors
+The autocomplete snippets are generated as follows:
+\begin{enumerate}
+	\item Parse the contents of the current file into an AST.
+	\item Find the node path corresponding to the cursor by traversing the AST.
+	\item Generate programs starting from the node path of the cursor.
+	\item Present the programs to the user as code completion snippets.
+\end{enumerate}
+
+\begin{figure}[ht]
+	\centering
+	\begin{tabular}{c c}
+		\begin{subfigure}[h]{6.25cm}
+			\centering
+			\includegraphics[width=1.00\textwidth]{vim1}
+			\caption{Inserting a for loop} \label{fig:vim1}
+		\end{subfigure}
+		\begin{subfigure}[h]{6cm}
+			\centering
+			\includegraphics[width=1.00\textwidth]{vim2}
+			\caption{Inserting an object property} \label{fig:vim2}
+		\end{subfigure}
+	\end{tabular}
+	\caption{Screenshots of vim using autocomplete of programs}
+\end{figure}
+
+The vim autocomplete feature is limited to inserting program snippets on a single
+line, and requires the user to explicitly trigger the \texttt{completefunc}
+feature to be prompted with code snippets. To further explore using our model
+for editing, we created a web-based editor that more deeply integrates Probabilistic
+Programming.
+
+\subsection{Guided IDE}
+
+The Guided IDE offers users an environment for building a program entirely using
+our model, using program snippets generated at the cursor's context as building
+blocks. The editor presents a window for the program text, and and list of
+suggestion code snippets. The user can move a cursor through the code window at
+positions corresponding to boundaries of program nodes. As the user moves the
+cursor, the snippet suggestions are regenerated using the cursor's node path, to
+continuosly provide suggestions relevant to the current context of the program. When the
+user selects a suggestion snippet, its AST is inserted into the program's AST at
+the cursor's position in the tree, and the code window is redrawn. The user can
+move the cursor within newly inserted blocks, so that the program can grow
+deeper. The initial corpus for the IDE's model consists of the JavaScript files
+that comprise our project.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.00\textwidth]{screenshot}
+  \caption{Screenshot of the web program editor GUI} \label{fig:screenshot}
+\end{figure}
 
@@
 
-\subsection{Limitations of the IDE}
-
-Usability issues
+We can use the Guided IDE to generate programs. The programs look like actual
+programs but do not actually function unless the user is very careful about
+which snippets they pick. Crucially, the variable and identifier names that our
+system generates are not in general meaningful in the scope of the generated
+program.  Whereas with the vim autocompletions the identifiers would often be
+valid because because the corpus was the current file, with the larger corpus of
+the Guided IDE and writing a program from scratch, the user is less likely to
+find program snippets that use valid identifiers.
 
@@
 
-Choosing valid identifiers, matching identifiers
+It is a limitation of the current model that identifier names are chosen
+probabilistically based on their position in the code, without regard to whether
+they are in scope and without refering to anything in another part of the
+program. Because of this limitation, the Guided IDE is not useful for practical
+purposes. We consider some alternative approaches for the Guided IDE using a
+similar model.
+
+\subsection{Alternative approaches for the IDE}
+
+\subsubsection{Prompt for identifiers}
+
+Rather than picking identifier names from the model, use placeholder identifiers
+and prompt the user to enter actual identifiers, with suggestions based on the
+current variables in scope and previously used identifiers. This would give the
+user the opportunity to create semantically valid programs, but the entering of
+identifier names could be tedious.
+
+\subsubsection{Stubs}
+
+Instead of having a movable cursor in the code window, have stubs in the
+program. Each stub would be a marker in place of an AST node. The user would
+select a stub and be given choices of snippets to replace it. Selecting a
+snippet would move the current snippet into the program in place of the selected
+stub.  The generated snippets could themselves have stubs in them, so that the
+user may continue to complete stubs and expand their program.  At the level of
+variable identifiers, the snippet suggestions would be taken from the computed
+list of in-scope variables at the position of the selected stub, unless the
+completion was for a variable declaration, in which case the user would be
+prompted to name a new variable.
+
+This approach, while offering less places to insert code then the current
+approach, would allow for programs to more closely be generated from our model,
+since the stub completions would always be at leaf nodes of the AST, whereas in
+the current approach the user may choose to insert a snippet in the middle of a
+list, which means that the transition from the inserted snippet to the following
+node in the list does not in general follow from the model.
+
+While the stub approach would be more restrictive for the user as it requires
+them to build the program in a certain way, with proper handling of identifiers,
+it may provide for an interesting or useful paradigm for interacting with a program
+source.
 
@@
 
-\clearpage
+%Other applications of the model: genetic algorithms
+
+\subsection{Extending the model}
+
+Just as our approach of modeling a program with a tree-based Markov chain is a
+higher order model than a linear Markov chain of the program source code tokens,
+there may be applications of probabilistic modeling to higher-order
+representations of programs than the AST, such as a higher-order abstract syntax
+\cite{hosa}. Such an application may model the relationship between different
+parts of an AST, and handle scoping of variables to model identifiers
+abstractly.
+
+With a more sophisticated model of identifiers that could generate programs with
+identifiers in meaningful relation to eachother, Probabilistic Programming could
+be used to generate semantically valid programs, for greater assistive,
+expressive, and programmatic power.
+
 \begin{thebibliography}{1}

### 2014-04-21 paper/paper.tex
Reduce margins
Add Sample AST
Add Source link
(Mostly) finish paper

@@
 \usepackage{listings}
+\usepackage[margin=3cm]{geometry}
 
-\usepackage{qtree}
-\usepackage{tikz}
-\usepackage{tikz-qtree}
+%\usepackage{qtree}
+%\usepackage{tikz}
+%\usepackage{tikz-qtree}
 
-\usepackage[]{algorithm2e}
+%\usepackage[]{algorithm2e}
 \hyphenation{ECMAScript JavaScript escodegen esprima}
@@
 in the list. If the statement is last in its container list, we connect its \emph{follow} edge to a special
-{\tt {\_end}} node.
+{\tt {\_end}} node. (See figure~\ref{fig:sample-ast} for an example of a AST.)
 
-\begin{center}
-
-[example of AST here]
+% \begin{center}
 
@@
 
-\end{center}
+% \end{center}
 
@@
 We use the syntax tree format of Mozilla's Parser API \cite{parser api}, and \emph{esprima} \cite{esprima}, an ECMAScript
+(the language standard of which JavaScript is an implementation)
 parser, to transform program source code into an AST of this format.
-
-%escodegen \cite{escodegen} (ECMAScript code generator) to transform ASTs into program source code.
 
@@
 To build our Markov chain structure, we parse a corpus of JavaScript files into their
-respective AST's, and then collect statistics from these trees to build our Markov model.
+respective AST's, and then collect statistics from these trees to establish probability distributions of nodes for a given AST context.
 
-The model maps AST paths to the probability distributions of node types or values for that given
-path. An AST path is a list of items alternating between a node type and one of its features, thus encoding the unique position of a 
-specific node in the AST. The probability distribution for an AST path is the set of possible node types or values that would appear in the
+An \emph{AST path} is a list of items, alternating between node types and features, that encodes the unique position of a 
+specific node in the AST. The \emph{probability distribution} for an AST path is the set of possible node types or values that would appear in the
 syntax tree following the given path, and the respective probabilities of their
-occurance following the given path.
+occurrence at that point. The model maps AST paths to the probability distributions of node types (or values, as we'll see later) for that given
+path.
 
 More formally, given a depth $d$ (indicating how many nodes we should backtrace) for a given
-${node}_d$ and feature ${edge}_d$, the generation of ${node}_d$'s child node given by the
-transition ${edge}_d$, can be expressed by the following function $f$, \begin{align*}
+${node}_d$ and feature ${edge}_d$, the generation of ${node}_d$'s child node given by
+ ${edge}_d$ can be expressed by the following function $f$: \begin{align*}
 f({node}_1, {edge}_1, ..., {node}_{d}, {edge}_{d}) = \begin{dcases*}
@@
 		{newnode_n},    &   $P({newnode_n})$ \\
-	\end{dcases*}
+	\end{dcases*},
 \end{align*}
 
-such that ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., and ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_k$ given the probabilities $P({newnode_1}), ..., P({newnode_n})$, and \begin{align*}
+where \begin{align*}
 \sum\limits_{i = 0}^n P({newnode_i}) = 1.
@@
 
-From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with, $\forall {edge} \in {Features}({newnode}_k)$, \begin{align*}
-{Path}({newnode}_k, {edge}) = f({node}_2, {edge}_2, ..., {node}_{d}, {edge}_{d}, {newnode}_k, {edge}).
+and ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_n$ given the probabilities \break $P({newnode_1}), ..., P({newnode_n})$.
+From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with the following formula.
+\begin{align*} &\forall {edge} \in {Features}({newnode}_k), \\
+& \hspace{3em} {Path}({newnode}_k, {edge}) = f({node}_2, {edge}_2, ..., {node}_{d}, {edge}_{d}, {newnode}_k, {edge}).
 \end{align*}
 
-This process will eventually terminate as the probability of generating {\tt \_end} nodes for a \emph{follow} edge will always be non-zero -- as the programs we parse into our corpus are surely finite -- and such nodes have no features to be ``expanded.''
+This process will eventually terminate as the probability of generating {\tt \_end} nodes for a \emph{follow} edge will always be non-zero -- the programs we parse into our corpus are surely finite -- and such nodes have no features to be ``expanded.''
 
@@
 Building the Markov chain model means being able to provide the probability distribution of new nodes for arbitrary input,
-as described by the function in Section 2.2. To do this, we constructed a data structure using JavaScript Object Notation (JSON),
+as described by the function in Section 2.2. To do this, we construct a data structure using JavaScript Object Notation (JSON),
 a structural format useful for ease of specification and integration with JavaScript programs.
@@
 The data structure is a hash table (also commonly called a map, dictionary, or associative array), that, on a given input, returns another hash table for paths that begin with that input.
-This behavior is repeated until the number of inputs reaches the desired depth, at which point a final hash table, mapping the node type to its probability distribution is returned.
+This behavior is repeated until the number of inputs reaches the desired depth, at which point a final hash table, mapping the node type to its probability distribution, is returned.
 This, in effect, is modularizing the function $f$ described above, so that we can reach the probability distribution with minimal code and data duplication.
-This is demonstrated by following example, for  ${depth} = 2$.
+This is demonstrated by following example, given  ${depth} = 2$.
 \begin{align*}
-lookup(lookup(lookup(lookup({model}, {node}_1), {edge}_1), {node}_2), {edge}_2) = \textbf{P}({nodes}).
+lookup(lookup(lookup(lookup({Model}, {node}_1), {edge}_1), {node}_2), {edge}_2) = \textbf{P}({newnodes}).
 \end{align*}
@@
 procedure addToModel(node, path):
-    model = corpusModel
+    model = CorpusModel
     for each elem in path[path.length-depth*2 .. path.length]:
@@
 
-When this process is complete, the JSON object is saved to a file and can be augmented by more code samples thereafter. (An example of such a model is shown in the Appendix.)
-
-See figure~\ref{fig:sample-model} in the appendix for a model for the program
-``var x = 34;" In this model we see for example that the path
-\texttt{[Program, body, VariableDeclaration, declarations]} leads to
-\texttt{VariableDeclarator} with probability 1. This indicates that a Program
-node whose body is a VariableDeclaration always has a declaration that is of
-type VariableDeclarator.
-
-The model is implemented as a JSON object where the path is indexed into the
-object by successively indexing by each string of the path.
+When this process is complete, the JSON object is saved to a file and can be augmented by more code samples thereafter.
+(An example of such a model is shown in figure~\ref{fig:sample-model} in the Appendix.
+In this model we see, for instance, that the path
+[\emph{Program, body, VariableDeclaration, declarations}] leads to
+\emph{VariableDeclarator} with probability 1. This indicates that a \emph{Program}
+node whose \emph{body} is a \emph{VariableDeclaration} always has a declaration that is of
+type \emph{VariableDeclarator}.)
 
@@
 
-\subsubsection{Random programs}
+Code can be generated rather easily using the approach explained in Section 2.2.
+The program tree is built up one node at a time, each one generated based on its unique AST path.
+The following algorithm demonstrates this process.
 
-Random programs can be generated rather easily using the approach explained in Section 2.2.
-The program tree is built up one node at a time, each one generated based on its unique AST path.
-We always begin a program with a ``Program'' node
+\begin{verbatim}
+procedure generateNode(model, path):
+    for each elem in path[path.length-depth*2 .. path.length]:
+        model = model[elem]
+    nodeType = pickRandomNodeType(model)
+    node = {`type':nodeType}
+    for each feature of nodeType:
+        new_path = path + [node.type, feature]
+        if feature is a list feature:
+            nodes = []
+            while (newNode = generateNode(model, new_path) != END_NODE:
+                nodes += [nodeNode]
+                new_path += [newNode.type, "FOLLOW"]
+            node.feature = nodes
+        else:
+            node.feature = generateNode(model, new_path)
+    return node
 
-\subsubsection{Snippets}
+generateNode(CorpusModel, [])
+\end{verbatim}
 
-% application of markov chain to abstact syntrax tree
+The algorithm looks very similar to the parse algorithm, as it is almost the same process, simply inverted.
+Rather than examine nodes, determine a path by building it up, and then increment a count in the model, this algorithm first finds the counts in the model for a \emph{given} path, generates a node, and generates nodes for all of that node's features recursively.
+The result is a generated AST for our program, which is then fed into \emph{escodegen} \cite{escodegen}, an ECMAScript code generator to output program source code that we can display and run.
+
+To generate random programs, we always begin with a ``Program'' node, and then add nodes to its \emph{body} feature (an array.) However, we quickly found that while entertaining, generating random programs had little potential for real application.
+Instead, we turned to generating code snippets (to be placed in specific contexts of the syntax tree.) Since our algorithm was general enough for this purpose, no more had to be done besides knowing the path for the node we wanted to generate.
+
+\subsubsection{Identifiers and literals}
+
+Naturally, identifiers and literals are a very large part of programming.
+Without them, code would be entirely unspecific, and would only consist of syntactic symbols and keywords.
+Yet they pose an issue with code generation, as we found generating identifiers or literals that are likely to be what the user is expecting an especially difficult problem to solve, let alone model.
+Struggling to come up with a simple, intelligent, and robust solution, and short on time, we fell back to using our existing Markov approach.
+
+In the current implementation of this system, following the paths [\emph{Identifier}, \emph{Name}], or [\emph{Literal}, \emph{Raw}] result not in node types, but rather in the text of the identifier names or literal values.
+This way, generating identifiers and literals is simply a probabilistic process like the rest of the code generation.
+Of course, this means that sometimes identifiers are added that are not in scope, that have mismatched types, or that simply make no semantic sense.
+(But adjusting $depth$ did seem to make programs more meaningful, expectedly.
+This can be seen experimentally on our web IDE, explained in Section 4.1.)
 
@@
 
-\subsection{vim autocomplete script}
+\subsection{\emph{vim} autocomplete script}
 
@@
 
+\subsection{Source}
+
+The source code for Probabilistic Programming is available online under an open
+source license:\\
+\url{https://github.com/RocHack/jschain}.\\
+The web IDE can be found here:
+\url{https://csug.rochester.edu/u/clehner/jschain/web/}.
+
 \begin{thebibliography}{1}
@@
 \begin{figure}[h!]
+	\caption{ast-sample.json}
+	\label{fig:sample-ast}
+	\centering
+	An AST generated from the program ``var x = 34;"
+
+	\lstinputlisting{ast-sample.json}
+\end{figure}
+
+\begin{figure}[h!]
 	\caption{model-sample.json}
@@
 	A model generated from the program ``var x = 34;"
+
+	\lstinputlisting{model-sample.json}
 \end{figure}
 
-\lstinputlisting{model-sample.json}
-
 \end{document}

### 2014-04-21 paper/paper.tex
Work on abstract and intro

@@
 \usepackage{listings}
-\usepackage[margin=3cm]{geometry}
+\usepackage[margin=3.75cm]{geometry}
 
@@
 \author{
-	Carlos Diaz
-	\and
-	Dan Hassin
-	\and
-	Charles Lehner
-	\and
-	Dan Viterise
+	Carlos Diaz, Dan Hassin, Charles Lehner, Dan Viterise \\
+    \vspace{0pt}\\ Department of Computer Science \\University of Rochester
 }
@@
 \begin{abstract}
-	We present Probabilistic Programming, a system for modeling abstract syntax
-	trees of programs with Markov chains, which we use to generate new programs.
-	We attempt to use this as an assistive device to guide programmers through
+	We present Probabilistic Programming, a system for generating JavaScript programs and code snippets based off a corpus of JavaScript source files.
+	In this paper, we will demonstrate storing statistical
+    data from non-linear structures (abstract syntax trees of programs) in a Markov model,
+    and of course using the model to generate new trees. We present a proof of concept for a web-based
+    assistive program editor that utilizes this technology to guide programmers through
 	constructing JavaScript programs.
@@
 
-\subsection{Motivation}
-
-For many people just starting out, programming can seem like a confusing and
-tedious activity. It can be unclear for many new programmers where they should
-begin and what should come next while coding. Without any proper guidance, a
-novice programmer is left on their own to scour thousands of pages online to
-discover what should come next in their code. Then, once a person becomes more
-adept at programming, he or she may notice that they are typing out the same or
-similar lines of code within each new program. This repetitiveness slows down
-the user and can make programming a much more tiresome task than it needs to be.
-These observations steered our group towards developing a way for someone to
-code with more instruction and productivity. The Probabilistic Programming
-method allows for a user to code in JavaScript using a web application. This web
-app provides a list of suggestions for the user's next line of code depending on
-where the user is in their program. Not only is this helpful for new programmers
-that need suggestions for what to write next in their code, this system allows
-for expert programmers to write code with more ease and efficiency.
+This paper will begin by discussing Markov chains and how we set out to use them for our goal.
+We will then describe in detail how we implemented a Markov chain for program syntax trees, how we were able to generate code using it, and finally we'll discuss our results and address the limitations of our product.
 
@@
 
-We chose to use a Markov chain to drive the code snippet generation, because
-(a) , and (b). A Markov Chain is a stochastic process
-that uses the current state and a set of probabilities to determine what state
-comes next. In the most simple case of a Markov Chain, the next action is based
-only upon the current situation and all other previous actions are forgotten.
-Thus, this process is essentially memoryless because all past states can be
-forgotten once a new state is reached. A good example of a Markov Chain is a
-random walk where at each integer value, the process chooses a new value such
-that the two possible values are the current value plus 1 and the current value
-minus 1. The transition probabilities for both of these options is .5. This is a
-good example of a Markov Chain because the next value on the random walk only
-depends on the current value in the walk; all previous values are not needed to
-choose the next value \cite{markov}.
+A Markov chain is a mathematical system that undergoes a stochastic process whereby
+the current state and a set of probabilities are used to determine what state
+to transition to next. In the most simple case of a Markov chain, the next action is based
+only upon the current state and all other previous actions are forgotten.
+Thus, this process is essentially ``memoryless'' because all past states can be
+forgotten once a new state is reached \cite{markov}. In this way, Markov chains can be likened
+to deterministic finite automata. So while they cannot remember a limitless amount of their history,
+calling them ``memoryless'' may be misleading, as they can indeed be implemented to remember a fixed
+size of their state history.
+
+Markov chains find application in many different fields.
+Familiar with Markov chains used for generating random human-sounding text (by training a model based on bigrams and trigrams of common speech), we were interested in perhaps more novel domains.
+We chose to use a Markov chain to drive code generation, because, like text, we feel that code can be personal and expressive, and the idea of generating code possibilities from a training set seemed like exciting research.
+It was also challenging -- since randomly generating code tokens would most likely produce syntactically invalid programs, we looked at program abstract syntax trees.
+And as the research developed, we found ourselves with a new vision of a semifeasable product.
+
+\subsection{Emerging goal}
+
+For many novice programmers, programming can seem like a confusing and
+tedious activity. It can be unclear for many new programmers where they should
+begin and what should come next while coding. They may conceptually understand
+what the next step should be, and yet be unable to express it with code. Or, they
+could be unsure of the right next step conceptually, and some code could inspire
+them to understand it.
+
+Even once one becomes more
+adept at programming, he or she may notice typing the same or
+similar lines of code in each new program. This repetitiveness slows down
+the user and can make programming more tiresome a task.
+
+The prospect of using a Markov chain to generate code snippets in an assistive, predictive, integrated development environment (IDE) for JavaScript seemed to be a natural progression from our research.
+Our final product, a Probabilistic Programming
+web application, allows a user to code in JavaScript by way of selecting from
+a list of suggestions for the user's next line of code. We imagine that this won't only be helpful for new programmers,
+but that this system would allow for expert programmers to write code with more ease and efficiency.
 
@@
 
-and ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_n$ given the probabilities \break $P({newnode_1}), ..., P({newnode_n})$.
+\noindent and ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_n$ given the probabilities $P({newnode_1}), ..., P({newnode_n})$.
 From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with the following formula.
@@
 The following algorithm demonstrates this process.
+
+\clearpage
 
@@
 our model, using program snippets generated at the cursor's context as building
-blocks. The editor presents a window for the program text, and and list of
+blocks. The editor presents a window for the program text, and a list of
 suggestion code snippets. The user can move a cursor through the code window at
@@
 cursor, the snippet suggestions are regenerated using the cursor's node path, to
-continuosly provide suggestions relevant to the current context of the program. When the
+continuously provide suggestions relevant to the current context of the program. When the
 user selects a suggestion snippet, its AST is inserted into the program's AST at
@@
 
-Instead of having a movable cursor in the code window, have stubs in the
+Instead of having a movable cursor in the code window, consider stubs in the
 program. Each stub would be a marker in place of an AST node. The user would
@@
 
-\section{Conclusions \& Open Problems}
+\section{Conclusions}
 
@@
 there may be applications of probabilistic modeling to higher-order
-representations of programs than the AST, such as a higher-order abstract syntax
-\cite{hosa}. Such an application may model the relationship between different
+representations of programs than the AST, such as a \emph{higher-order abstract syntax}, which promises to ``incorporate name binding information in a uniform and language generic way'' \cite{hosa}.
+Such an application may model the relationship between different
 parts of an AST, and handle scoping of variables to model identifiers
@@
 With a more sophisticated model of identifiers that could generate programs with
-identifiers in meaningful relation to eachother, Probabilistic Programming could
+identifiers in meaningful relation to each other, Probabilistic Programming could
 be used to generate semantically valid programs, for greater assistive,
 expressive, and programmatic power.
+
+\subsection{Future}
+
+We should
 
@@
 The source code for Probabilistic Programming is available online under an open
-source license:\\
-\url{https://github.com/RocHack/jschain}.\\
-The web IDE can be found here:
+source license:
+\url{https://github.com/RocHack/jschain}.
+
+The web IDE can be found at
 \url{https://csug.rochester.edu/u/clehner/jschain/web/}.

### 2014-04-21 paper/paper.tex
Add more identifier ideas

@@
 
+We explored several other approaches that could solve the issue of identifiers and literals.
+
 \subsubsection{Prompt for identifiers}
@@
 
-\section{Conclusions}
+\subsubsection{Secondary models for identifiers}
 
-%Other applications of the model: genetic algorithms
+When generating identifiers, we were interested in generating meaningful ones rather than random ones (whose weights were calculated from counts originating from identifiers that could be in entirely different places of the program.) We wanted to design a model that could keep track of when the same variables were used in the same (or similar) place.
 
-\subsection{Extending the model}
+One idea was to, during the parse, keep track of the path at which each variable identifier was used, grouped by the identifier name.
+When an identifier needs to be generated, we can find that path in our secondary model, and if so, bind to a random (but weighted) identifier that has been used at that path, with the weight being based on the frequency of that identifier appearing at that path.
+For subsequent statements that need identifiers, the same process could occur, and, if the same variable group was chosen, it could generate the variable name that was bound to the first statement.
+
+Another idea also involved constructing a secondary model during the parse, that kept track of statements that used the same identifier.
+When such statements are found, it would compute the path between those statements and increment a counter associated that path.
+This would represent the likelihood of two nodes connected by that path using the same identifier.
+Then, during code generation, whenever an identifier is needed, the generator would make a list of all variables in scope at that point, and compute the path from the variable declaration (and/or last use) to the node being generated.
+It would look up each path in the secondary model and assign each of the in-scope variables a weight.
+A variable could then be chosen based on this probability.
+
+We suspect that this may affect the complexity of our algorithms significantly.
+If what the system is doing now is analogous to the behavior of a context-free grammar (i.e.
+by recursively generating nodes), having detection of bound (same) identifiers across nodes would involve pattern-matching and, perhaps even analogy-solving, could be considered more of an action of a Turing machine.
+This could mean the complexity of the process could be unbounded.
+
+Unfortunately, we were unable to implement these ideas or explore them further due to time constraints.
+
+\subsubsection{Extending the model}
 
@@
 
-\subsection{Future}
+\section{Conclusions}
 
-We should
+We consider the web IDE a strong proof of concept for context-based code snippet generation.
+We are aware of some of its serious structural limitations, but we are confident that related concepts, such as genetic algorithms, might be possible.
+Specifically, we are interested in the possibility of creating genetic algorithms that take advantage of this platform to build and ``evolve'' code in a similar process.
+Perhaps it could even be given a purpose; for instance, to satisfy a set of given test cases.
 
-\subsection{Source}
+%Other applications of the model: genetic algorithms
+
+\subsection*{Source \& Demo}
 
@@
 source license:
-\url{https://github.com/RocHack/jschain}.
+\url{https://github.com/RocHack/jschain}.\\
 
-The web IDE can be found at
+\noindent The web IDE can be found at
 \url{https://csug.rochester.edu/u/clehner/jschain/web/}.
+
+\clearpage
 

### 2014-04-21 build-corpus.js
Generate corpusi of different depths

@@ -3,36 +3,51 @@ var esprima = require("esprima");
 var parse = require("./parse");
 
 var modelDest = "web/corpus-model.js";
-var model = {};
 
 var space = +process.argv[2] || 0;
-var depth = 4;
 
-var filesDir = "js-files";
-fs.readdirSync(filesDir).forEach(function (name) {
-	if (!name.match(/\.js$/)) return;
-	var source = fs.readFileSync(filesDir + "/" + name).toString();
-	if (source[0] == "#") {
-		// Remove shebang which esprima doesn't like
-		source = source.substr(source.indexOf("\n")+1);
-	}
-	console.log("Parsing", name, "(" + source.length + " bytes)");
-	var syntax;
-	try {
-		syntax = esprima.parse(source);
-	} catch(e) {
-		console.error("Error parsing JS:", e);
-		return;
-	}
-	try {
-		parse.parseSyntax(syntax, null, depth, model);
-	} catch(e) {
-		console.error("Error parsing syntax:", e);
-	}
-});
 
-console.log("Writing model to", modelDest, space ? "with space=" + space : "");
-var corpusJSON = JSON.stringify(model, null, space);
-var corpusOutput = "window.corpusModel = " + corpusJSON;
-fs.writeFileSync(modelDest, corpusOutput);
-console.log(corpusOutput.length + " bytes written to " + modelDest);
+function buildCorpus(depth) {
+
+	var model = {};
+
+	var filesDir = "js-files";
+	fs.readdirSync(filesDir).forEach(function (name) {
+		if (!name.match(/\.js$/)) return;
+		var source = fs.readFileSync(filesDir + "/" + name).toString();
+		if (source[0] == "#") {
+			// Remove shebang which esprima doesn't like
+			source = source.substr(source.indexOf("\n")+1);
+		}
+		console.log("Parsing", name, "(" + source.length + " bytes)");
+		var syntax;
+		try {
+			syntax = esprima.parse(source);
+		} catch(e) {
+			console.error("Error parsing JS:", e);
+			return;
+		}
+		try {
+			parse.parseSyntax(syntax, null, depth, model);
+		} catch(e) {
+			console.error("Error parsing syntax:", e);
+		}
+	});
+
+	console.log("Writing model to", modelDest, space ? "with space=" + space : "");
+	var corpusJSON = JSON.stringify(model, null, space);
[…]

### 2014-04-21 web/main.js

@@ -12,9 +12,16 @@ var escodegenOptions = {
 	}
 };
 var END = "_end";
-var depth = 4;
 
-var model = window.corpusModel;
+var depth;
+var model;
+
+window.setDepth = function (d) {
+	depth = d;
+	model = window.corpusModel[d];
+}
+
+setDepth(4);
 
 var tree;
 
@@ -93,23 +100,14 @@ function setCurrentPath(path) {
 
 window.setCurrentPathToNode = function (node) {
 
-	console.log("setting current path to node ",node);
-
 	if (node.type == "ExpressionStatement")
 		node = node.expression;
 
 	// console.log("looking for", node, " in ",tree);
 	currentPosition = getPath(tree, node);
-	if (!currentPosition) {
-		console.log("tree:", JSON.stringify(tree), 'Node:', JSON.stringify(node));
-	}
-
-	console.log("current position is ",currentPosition);
 }
 
 window.insertSnippet = function (node) {
-
-	console.log("inserting snippet", node);
 
 	if (!tree)
 	{
@@ -142,7 +140,7 @@ function labelNodeId(node) {
 window.getNodeById = function (id) {
 	var node = nodesById[id];
 	if (!node) {
-		console.log("no node found with id", id);
+		//console.log("no node found with id", id);
 	}
 	return node;
 };
@@ -164,7 +162,7 @@ function generateProgramSource() {
 }
 
 window.getSnippets = function (num) {
-	console.log("**** generating snippet from path ",currentPosition.path);
+	console.log("path =",currentPosition.path);
 	return array(num).map(generateProgramSource).filter(Boolean);
 };
 
@@ -172,17 +170,14 @@ window.getCurrentPosition = function () {
 	return currentPosition;
 };
 
-window.deleteSnippet = function (node) {
-
-	console.log("deleting snippet", node);
-
+window.deleteCurrentSnippet = function () {
 	if (isNaN(currentPosition.idx)) 
 	{
 		currentPosition.container[currentPosition.idx] = null;
 	} 
 	else
 	{
-		currentPosition.container.splice(currentPosition.idx, currentPosition.idx);
+		currentPosition.container.splice(currentPosition.idx, 1);
 	}
 
 	//console.log("generating tree", tree);

### 2014-04-21 web/ui.js

@@ -49,7 +49,7 @@ function checkKey(e)
 	        // up arrow
 			currentSelection -= 1;
 	    }
-	    else if (e.keyCode == 40 && currentSelection < numSelections)
+	    else if (e.keyCode == 40 && currentSelection < numSelections - 1)
 	    {
 	        // down arrow
 	    	currentSelection += 1;
@@ -100,9 +100,9 @@ function moveCursor(forward) {
 	var node, elem, found;
 	function findStatement(i, el) {
 		var n = window.getNodeById($(el).data('node-id'));
-		console.log("NODE!", n, el.innerText);
+		// console.log("NODE!", n, el.innerText);
 		if (n) {
-			console.log("go to children");
+			// console.log("go to children");
 			$(el).children('.statement').eachDir(findStatement, forward);
 			if (found) return false;
 		}
@@ -120,13 +120,13 @@ function moveCursor(forward) {
 	while (!found && from.length) {
 		if (i++ > 50) throw new Error("too much recursion");
 		from[dirAll]('.statement').each(findStatement);
-		console.log("go to parent", from);
+		// console.log("go to parent", from);
 		from = from.parent();
 	}
 	if (!from) {
 		// set it to the beginning
 	}
-	console.log("Found:", found, "elem", elem && elem.innerText);
+	// console.log("Found:", found, "elem", elem && elem.innerText);
 	if (!node) return;
 	window.setCurrentPathToNode(node);
 	cursor.insertAfter(elem);
@@ -187,13 +187,22 @@ function restoreCursor(node)
 
 function deleteKey()
 {
-	var newCode = window.deleteSnippet(cursor.prev());
+	var cursorPos = cursor.prev().prev();
+	var nodeid = $(cursorPos).data('node-id');
+
+	console.log('cursorPos = ',cursorPos[0], 'id = ',nodeid);
+
+	console.log('getting node by id '+nodeid);
+	var node = window.getNodeById(nodeid);
+
+	console.log('node = ',node);
+
+	var newCode = window.deleteCurrentSnippet();
 
 	$('#editor').html(newCode);
 
-	window.setCurrentPathToNode(cursor.prev());
-
-	restoreCursor(cursor.prev());
+	window.setCurrentPathToNode(node);
+	restoreCursor(nodeid);
 
 	newOptions();
 }

### 2014-04-21 web/main.js
Add UI for changing depths

@@ -17,6 +17,7 @@ var depth;
 var model;
 
 window.setDepth = function (d) {
+	console.log("switching to depth",d);
 	depth = d;
 	model = window.corpusModel[d];
 }

### 2014-04-21 web/ui.js

@@ -134,6 +134,17 @@ function moveCursor(forward) {
 	newOptions();
 }
 
+function selectDepth(li) {
+	var depth = $(li).text();
+	if (depth == $('#current-depth').html())
+		return;
+
+	$('#current-depth').html(depth);
+	window.setDepth(+depth);
+
+	newOptions();
+}
+
 function scrollIntoView(container, elem) {
 	var offset = cursor.offset().top - editor.offset().top;
 	if (offset < 0 || offset > container.height()) {
@@ -232,6 +243,8 @@ $(document).ready(function() {
 	cursor = $('#cursor');
 	addHandlers();
 	newOptions();
+
+	$('.dropdown-menu').dropdown();
 
 	$('#output').hide();
 	$(".run-btn").click(function () {

### 2014-04-21 web/main.js
Display current path in header

@@ -163,6 +163,13 @@ function generateProgramSource() {
 }
 
 window.getSnippets = function (num) {
+	if (currentPosition.path.length > 0)
+	{
+	    var path = currentPosition.path.slice(-(depth)*2);
+
+		$('h1').css('font-size','1.75em');
+		$('h1').text(path.join(", "))
+	}
 	console.log("path =",currentPosition.path);
 	return array(num).map(generateProgramSource).filter(Boolean);
 };

### 2014-04-21 web/ui.js
Syntax highlighting

@@ -179,6 +179,8 @@ function complete(option)
 
 	$('#editor').html(newCode);
 
+$('#editor').each(function(i, e) {hljs.highlightBlock(e)});
+
 	if (node.type != 'Program')
 		window.setCurrentPathToNode(node);
 
@@ -212,6 +214,7 @@ function deleteKey()
 
 	$('#editor').html(newCode);
 
+
 	window.setCurrentPathToNode(node);
 	restoreCursor(nodeid);
 
@@ -227,6 +230,7 @@ function newOptions()
 	{
 		code = snips[i];
 		var pre = $('<pre></pre>').html(code);
+		pre.each(function(i, e) {hljs.highlightBlock(e)});
 		$('#options').append($('<li></li>').append(pre));
 	}
 

### 2014-04-21 paper/paper.tex
New screenshow to show syntax highlighting and depth toggle

@@
 Of course, this means that sometimes identifiers are added that are not in scope, that have mismatched types, or that simply make no semantic sense.
-(But adjusting $depth$ did seem to make programs more meaningful, expectedly.
-This can be seen experimentally on our web IDE, explained in Section 4.1.)
+(But adjusting $depth$ did seem to make programs more meaningful, as expected.
+This can be seen experimentally by modifying depth on our web IDE, which explained in Section 4.1.)
 
@@
   \centering
-  \includegraphics[width=1.00\textwidth]{screenshot}
+  \includegraphics[width=1.00\textwidth]{ide}
   \caption{Screenshot of the web program editor GUI} \label{fig:screenshot}

### 2014-04-21 paper/paper.tex
Small paper edits

@@
 \usepackage{listings}
+\usepackage[labelfont=bf]{caption}
 \usepackage[margin=3.75cm]{geometry}
@@
 (But adjusting $depth$ did seem to make programs more meaningful, as expected.
-This can be seen experimentally by modifying depth on our web IDE, which explained in Section 4.1.)
+This can be seen experimentally by modifying depth on our web IDE, which is discussed in Section 4.1.)
 
@@
 suggestions of programs upon the user's request, using vim's
-\texttt{completefunc} feature. When the user is editing a JS file in vim with
+\texttt{completefunc} feature. When the user is editing a JavaScript file in vim with
 this feature, they may press a keyboard command to get a list of program
@@
 program path at cursor's position. The user may pick one of the snippet
-suggestions to have it inserted into their code at their cursor.
-
-The autocomplete snippets are generated as follows:
+suggestions to have it inserted into their code at their cursor. The autocomplete snippets are generated as follows:
 \begin{enumerate}
@@
 			\includegraphics[width=1.00\textwidth]{vim1}
-			\caption{Inserting a for loop} \label{fig:vim1}
+			\caption{Inserting a {\tt for} loop} \label{fig:vim1}
 		\end{subfigure}
@@
   \includegraphics[width=1.00\textwidth]{ide}
-  \caption{Screenshot of the web program editor GUI} \label{fig:screenshot}
+  \caption{Screenshot of the web program editor GUI.
+The header along the top indicates the path (trimmed to the selected depth) to the current position of the cursor.
+Code snippets on the left can be navigated using the arrow keys.} \label{fig:screenshot}
 \end{figure}
@@
 \subsection{Alternative approaches for the IDE}
-
-We explored several other approaches that could solve the issue of identifiers and literals.
 
@@
 
-\subsubsection{Secondary models for identifiers}
+\subsection{Intelligent handling of identifiers and literals}
+
+We explored other approaches to solve the issue of generating meaningful identifiers and literals described before.
+
+\subsubsection{Secondary models}
 
@@
 If what the system is doing now is analogous to the behavior of a context-free grammar (i.e.
-by recursively generating nodes), having detection of bound (same) identifiers across nodes would involve pattern-matching and, perhaps even analogy-solving, could be considered more of an action of a Turing machine.
-This could mean the complexity of the process could be unbounded.
+by recursively generating nodes), then detection of use of the same identifiers across nodes would involve pattern-matching and perhaps even analogy-solving.
+This could analogously be considered more like the action of a Turing machine, which would mean that the complexity of the process could be unbounded.
 
@@
 
-\subsubsection{Extending the model}
+\subsubsection{Using a higher-order model}
 
@@
 
-\section{Conclusions}
+\section{Conclusion}
 
@@
 
-\subsection*{Source \& Demo}
+\subsection*{Source code \& demo}
 
@@
 \begin{figure}[h!]
-	\caption{ast-sample.json}
+	\caption{JSON AST sample}
 	\label{fig:sample-ast}
 	\centering
-	An AST generated from the program ``var x = 34;"
+    \vspace{-7pt}
+	An AST generated from the program ``{\tt var x = 34;}"
 
@@
 \begin{figure}[h!]
-	\caption{model-sample.json}
+	\caption{JSON Markov model sample}
 	\label{fig:sample-model}
 	\centering
-	A model generated from the program ``var x = 34;"
+    \vspace{-7pt}
+	A Markov chain model generated from the program ``{\tt var x = 34;}"
 
